### 2013-10-23 plan.tex
Started an outline and added a single reference.

\section{Plan}

We propose to create a verified compiler for a language similar to that
discussed in Pottier and Simonet's work\cite{InfoFlowML}.  We will
construct this compiler using the Coq proof assistant\cite{coq?} using
features which allow us to extract a certified compiler.  This will
allow us to take the first step towards a verified compiler for a
language with static information flow types such as Jif\cite{Jif?} or
FlowCaml\cite{FlowCaml?}.

%TODO:
Stuff about why it would be useful to have a verified compiler for
something like Jif or FlowCaml.

%TODO:
Stuff about the target and source languages we will be supporting.

%TODO:
Expected timeline, anticipated problems, and potential additions if we
have time?

### 2013-10-24 plan.tex
Tweaks.

@@
 We propose to create a verified compiler for a language similar to that
-discussed in Pottier and Simonet's work\cite{InfoFlowML}.  We will
-construct this compiler using the Coq proof assistant\cite{coq?} using
-features which allow us to extract a certified compiler.  This will
-allow us to take the first step towards a verified compiler for a
-language with static information flow types such as Jif\cite{Jif?} or
-FlowCaml\cite{FlowCaml?}.
+discussed in Pottier and Simonet's work\cite{InfoFlowML} compiling to a
+simple register transfer level (RTL) language.  We will construct this
+compiler using the Coq proof assistant\cite{coq?} using features which
+allow us to extract a certified compiler.  This will allow us to take
+the first step towards a verified compiler for a language with static
+information flow types such as Jif\cite{myers1999jflow} or
+FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.
 
@@
 %TODO:
-Stuff about the target and source languages we will be supporting.
-
-%TODO:
 Expected timeline, anticipated problems, and potential additions if we

### 2013-10-24 plan.tex
Draft of plan.  Added some references.

@@
 simple register transfer level (RTL) language.  We will construct this
-compiler using the Coq proof assistant\cite{coq?} using features which
-allow us to extract a certified compiler.  This will allow us to take
-the first step towards a verified compiler for a language with static
-information flow types such as Jif\cite{myers1999jflow} or
-FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.
+compiler using the Coq proof assistant\cite{coqart} using features which
+allow us to extract a certified compiler.
 
-%TODO:
-Stuff about why it would be useful to have a verified compiler for
-something like Jif or FlowCaml.
+This will allow us to take the first step towards a verified compiler
+for a language with static information flow types such as
+Jif\cite{myers1999jflow} or
+FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.  Developing a
+verified compiler for such languages removes the need for programmers to
+trust the compiler, they can instead verify the proofs used to generate
+the compiler to convince themselves that semantics of the original
+source program are preserved and properties such as non-interference are
+enforced by the compiler.  This rules out the possibility that a
+compiler does not introduce bugs not present in the source program, such
+as additional undesirable information flows.  As a result, this compiler
+reduces the trusted computing base from a large complex compiler to a
+much simpler verifier to mechanically check the proofs the compiler is
+extracted from.
 
-%TODO:
-Expected timeline, anticipated problems, and potential additions if we
-have time?
+To develop this verified compiler using Coq, we plan to write formal
+proofs to show that the type system correctly enforces non-interference
+and that a CPS conversion\cite{appel2007compiling}, which would be our
+approach to compiling the language to an RTL language, correctly
+preserves the semantics of the source program.  We expect that the proof
+that the type system enforces non-interference will be very similar to
+the proof presented in Pottier and Simonet's paper\cite{InfoFlowML}.
+Development will proceed by first writing the proof for CPS conversion,
+developing a compiler for the language without enforcing the type
+system, and then going back and adding the non-interference proof.

### 2013-10-24 background.tex
more background, a couple more references

\section{Background}

Much work has been completed on information-flow languages.  Projects like FlowCaml\cite{simonet2003flow}, JIF\cite{myers2000protecting}, Fabric\cite{liu2009fabric}, and more offer complete production languages with strong information-flow security guarantees.  These languages are aimed for deployment in industry, where their security properties are relied upon for the integrity and confidentiality of sensitive information.  Like all software systems, the compilers for these languages are not free of bugs; these bugs can compromise the very security policies that the languages seek to enforce.  Other languages have addressed this problem by building verified compilers\cite{leroy2012compcert,okuma2003executing,chlipala2010verified,berghofer2004extracting,strecker2002formal,necula1998design,necula2002proof}.  Verified compilers are free of implementation-level bugs; they guarantee that the semantics of the source program are identical to the semantics of the destination language\cite{necula1998design}.  As of yet, no verified compiler exists for an information-flow language, leaving users to choose between a provably correct compiler for an insecure language and a potentially-buggy compiler for a secure language.

%Both verified compilers and compilers for information-flow languages are nothing new.  (most of the citations in the bibtex file could apply here).  Much previous work has been done on verified compilers, notably work on CompCert.  Much work has also been done in information-flow languages.  Information flow languages promise statically-enforced security properties, and often have proofs demonstrating that the language is capable of enforcing these policies.  But how do you trust the language?  Discussion of motivating bugs!  Verified compilers are a way to avoid this problem entirely.  A fun Fabric bug was a failure to support the += operator, which was silently accepted and treated as =, which is terrible.  But it doesn't actually affect the security aspect of these languages.  Anyway.

### 2013-10-24 plan.tex
grammar correction

--- 2013-10-24 plan.tex
@@
 
-We propose to create a verified compiler for a language similar to that
-discussed in Pottier and Simonet's work\cite{InfoFlowML} compiling to a
-simple register transfer level (RTL) language.  We will construct this
-compiler using the Coq proof assistant\cite{coq?} using features which
-allow us to extract a certified compiler.  This will allow us to take
-the first step towards a verified compiler for a language with static
-information flow types such as Jif\cite{myers1999jflow} or
-FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.
+We propose a verified compiler for a language similar to typed CoreML$^2$\
+cite{InfoFlowML}, targeting a simple register transfer level (RTL) language.  
+We will construct this compiler in the Coq proof assistant\cite{coqart}, 
+which allows extraction to OCAML\cite{berghofer2004extracting}.
 
-%TODO:
-Stuff about why it would be useful to have a verified compiler for
-something like Jif or FlowCaml.
+This will allow us to take the first step towards a verified compiler
+for a language with static information flow types such as
+Jif\cite{myers1999jflow} or
+FlowCaml\cite{simonet2003flow,simonet2003flowmanual}.  Developing a
+verified compiler for such languages removes the need for programmers to
+trust the compiler; to convince themselves that the compiler preserves 
+the semantics and enforces properties (such as non-interference) of the 
+source program they can verify the proofs used to generate
+the compiler.  This eliminates the possibility of a compiler
+introducing bugs during the translation of the source program, such
+as additional undesirable information flows.  As a result, this compiler
+reduces the trusted computing base from a large complex compiler to a
+much simpler verifier to mechanically check the proofs from which 
+the compiler is extracted.
 
-%TODO:
-Expected timeline, anticipated problems, and potential additions if we
-have time?
+To develop this verified compiler using Coq we plan to write formal
+proofs which show that the type system correctly enforces non-interference
+and that a CPS conversion\cite{appel2007compiling}, which would be our
+approach to compiling the language to an RTL language, correctly
+preserves the semantics of the source program.  We expect the proof
+that the type system enforces non-interference will be very similar to
+the proof presented in Pottier and Simonet's paper\cite{InfoFlowML}.
+Development will proceed by first writing the proof for CPS conversion,
+developing a compiler for the language without enforcing the type
+system, and then going back and adding the non-interference proof.

### 2013-10-24 plan.tex
cleaned up phrasing

@@
 
-We propose a verified compiler for a language similar to typed CoreML$^2$\
-cite{InfoFlowML}, targeting a simple register transfer level (RTL) language.  
+We propose a verified compiler for a language similar to typed CoreML$^2$
+\cite{InfoFlowML}, targeting a simple register transfer level (RTL) language.  
 We will construct this compiler in the Coq proof assistant\cite{coqart}, 
-which allows extraction to OCAML\cite{berghofer2004extracting}.
+which allows the extraction of verified code to OCAML\cite{berghofer2004extracting}.
 
@@
 verified compiler for such languages removes the need for programmers to
-trust the compiler; to convince themselves that the compiler preserves 
-the semantics and enforces properties (such as non-interference) of the 
-source program they can verify the proofs used to generate
-the compiler.  This eliminates the possibility of a compiler
+trust the compiler; users can convince themselves of the compiler's 
+correctness by verifying the proofs used to generate the compiler.
+This eliminates the possibility of a compiler
 introducing bugs during the translation of the source program, such
@@
 system, and then going back and adding the non-interference proof.
+We plan to target CompCert's RTL intermediate language.

### 2013-12-17 proposal/background.tex
Moving stuff because I'm going insane with everything in one big directory.

--- 2013-10-24 background.tex
@@
 
-Much work has been completed on information-flow languages.
-Projects like FlowCaml\cite{simonet2003flow}, JIF\cite{myers2000protecting}, Fabric\cite{liu2009fabric}, and more offer complete production languages with strong information-flow security guarantees.
-These languages are aimed for deployment in industry, where their security properties are relied upon for the integrity and confidentiality of sensitive information.
-Like all software systems, the compilers for these languages are not free of bugs; these bugs can compromise the very security policies that the languages seek to enforce.
-Other languages have addressed this problem by building verified compilers\cite{leroy2012compcert,okuma2003executing,chlipala2010verified,berghofer2004extracting,strecker2002formal,necula1998design,necula2002proof}.
-Verified compilers are free of implementation-level bugs; they guarantee that the semantics of the source program are identical to the semantics of the destination language\cite{necula1998design}.
-As of yet, no verified compiler exists for an information-flow language, leaving users to choose between a provably correct compiler for an insecure language and a potentially-buggy compiler for a secure language.
+Much work has been completed on information-flow languages.  Projects
+like FlowCaml\cite{simonet2003flow}, JIF\cite{myers2000protecting},
+Fabric\cite{liu2009fabric}, and more offer complete production languages
+with strong information-flow security guarantees.  These languages are
+aimed for deployment in industry, where their security properties are
+relied upon for the integrity and confidentiality of sensitive
+information.  Like all software systems, the compilers for these
+languages are not free of bugs; these bugs can compromise the very
+security policies that the languages seek to enforce.  Other languages
+have addressed this problem by building verified
+compilers\cite{leroy2012compcert,okuma2003executing,chlipala2010verified,berghofer2004extracting,strecker2002formal,necula1998design,necula2002proof}.
+Verified compilers are free of implementation-level bugs; they guarantee
+that the semantics of the source program are identical to the semantics
+of the destination language\cite{necula1998design}.  As far as we know,
+no verified compiler exists for an information-flow language, leaving
+users to choose between a provably correct compiler for an insecure
+language and a potentially-buggy compiler for a secure language.
 

### 2013-12-17 paper/paper.tex
Really rough start to an outline of a paper.

\documentclass[a4paper,twocolumn]{article}

\title{Towards a Mechanized Proof of Noninterference}
\author{Tom Magrino \and Mathew Milano}
\begin{document}

\maketitle

\abstract
Abstract goes here, summarize what this paper covers and what we ended up with.

\section{Introduction}
Introduce idea and outline what it will require us to do.  Explain where we're
starting from (Pottier and Simonet) at a high level.  Justify why this is an
interesting problem that needs tackling.

\section{Operational Semantics}
Outline the operational semantics of the baby language we developed.  Outline
any interesting similarities or differences between us and Pottier and Simonet.

\section{Type System}
Outline the type system of the language and again compare/contrast with
Pottier and Simonet.

\section{Proofs}

I'm imagining this section could outline the human version of the proofs we'd
need to do a full noninterference proof.

\section{Mechanizing Proofs}

Talk about what did/should have been mechanized.  Talk about troubles and/or
interesting bits here.  Might point out here that we're fairly certain that we
found an inaccuracy in the original paper.

\section{Related Work}

Discuss various references related to our project.

\section{Future Work}
Talk about stuff we still have to do for this to be a complete project.  Talk
about expected problems and required techniques.  Talk about what we could do
once this project is completed and we could use it for other projects.

\section{Conclusion}
Summarize again what we covered, where we ended up, and where we'd like to go
next.

\bibliographystyle{abbrv}
\bibliography{proposal.bib}
\section*{Acknowledgements}

Acknowledge people who really helped.  Primary name that comes to mind is that
nice Jon fellow we spent many hours on the phone getting advice from.
\end{document}

### 2013-12-17 paper/paper.tex
Whoops, fixed bibtex reference.

@@
 \section{Introduction}
+
 Introduce idea and outline what it will require us to do.  Explain where we're
@@
 \section{Operational Semantics}
+
 Outline the operational semantics of the baby language we developed.  Outline
@@
 \section{Type System}
+
 Outline the type system of the language and again compare/contrast with
@@
 \section{Future Work}
+
 Talk about stuff we still have to do for this to be a complete project.  Talk
@@
 \section{Conclusion}
+
 Summarize again what we covered, where we ended up, and where we'd like to go
@@
 
-\bibliographystyle{abbrv}
-\bibliography{proposal.bib}
 \section*{Acknowledgements}
@@
 nice Jon fellow we spent many hours on the phone getting advice from.
+
+\bibliographystyle{abbrv}
+\bibliography{paper.bib}
+
 \end{document}

### 2013-12-18 paper/paper.tex
Totally had matthew's name mispelled :P
Typing judgements for values.  Fixed some other small bits.

@@
 \documentclass[a4paper,twocolumn]{article}
+\usepackage{mathpartir}
+\usepackage{amssymb}
+\usepackage{float}
+\floatstyle{boxed}
+\restylefloat{figure}
+
+\newcommand{\langName}[0]{TinyML\textsuperscript{2}}
+\newcommand{\typeRule}[3]
+{#1 \textsf{ } \vdash \textsf{ } #2 \textsf{ } \colon \textsf{ } #3}
 
 \title{Towards a Mechanized Proof of Noninterference}
-\author{Tom Magrino \and Mathew Milano}
+\author{Tom Magrino \and Matthew Milano}
 \begin{document}
@@
 \abstract
+
+Citing something for now to make an error message go away~\cite{Pierce:SF}.
 Abstract goes here, summarize what this paper covers and what we ended up with.
@@
 
-Outline the type system of the language and again compare/contrast with
-Pottier and Simonet.
+\begin{figure}
+  \begin{mathpar}
+    \infer*[left=V-Int]
+    { }
+    {\typeRule{\Gamma}{k}{\textsf{int}^{*}}}
+
+    \infer*[left=V-Var]
+    {\Gamma(x) = t}
+    {\typeRule{\Gamma}{x}{t}}
+
+    \infer*[left=V-Abs]
+    {\typeRule{pc, \Gamma[x \mapsto t'][f \mapsto (t \to t')^l]}{e}{t}}
+    {\typeRule{\Gamma}{\textsf{fix} f . \lambda x . e}{(t \to t')^l}}
+
+    \infer*[left=V-Bracket]
+    {
+    \typeRule{\Gamma}{v_1}{t} \\
+    \typeRule{\Gamma}{v_2}{t} \\
+    l \in H \\
+    l \lhd t
+    }
+    {\typeRule{\Gamma}{<v_1 | v_2>}{t}}
+
+    \infer*[left=V-Sub]
+    {
+    \typeRule{\Gamma}{v}{t'} \\
+    t' \leq t
+    }
+    {\typeRule{\Gamma}{v}{t}}
+  \end{mathpar}
+  \caption{Typing judgements for values in \langName.}
+\end{figure}
+
+The type system for our simple language is given in Figure~\ref{fig:type_rules}.
 
@@
 
-Acknowledge people who really helped.  Primary name that comes to mind is that
-nice Jon fellow we spent many hours on the phone getting advice from.
+We would like to thank Jon Sailor from Facebook for helping us over the phone
+work through some of the more difficult problems we ran into using Coq.
 
 \bibliographystyle{abbrv}
-\bibliography{paper.bib}
+\bibliography{paper}
 

### 2013-12-18 paper/paper.tex
Added initial figures for typing rules.  Still need to fill out guards and subtyping.

@@
-\documentclass[a4paper,twocolumn]{article}
+\documentclass[a4paper]{article}
 \usepackage{mathpartir}
 \usepackage{amssymb}
-\usepackage{float}
-\floatstyle{boxed}
-\restylefloat{figure}
+%\usepackage{float}
+%\floatstyle{boxed}
+%\restylefloat{figure}
 
@@
 {#1 \textsf{ } \vdash \textsf{ } #2 \textsf{ } \colon \textsf{ } #3}
+\newcommand{\guards}[0]{\lhd}
 
@@
 
+\langName has a very simple type system, including only integers and lambda
+abstractions, which include fixpoints by default, along with appropriate
+security labels for both.  In \langName, security labels specify confidentiality
+policies, specifying what other data the labelled data can affect.\footnote{It is
+not difficult to imagine how to extend our system to include integrity labels.
+However, for the sake of simplicity, we decided to concern ourselves with only
+confidentiality at first}  Data can only be affected by items at the same label
+or lower (so high data can not affect low labelled data).
+
+The language's type system includes subtyping, which allows data that have a
+lower label to flow into higher labelled data.  The rules for subtyping are
+given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
+integers are subtypes of integers with equivalent or higher labels in the label
+lattice.  Similarly, rule \textsc{S-Abs} gives the standard subtyping
+relationship for functions along with the same label relation.
+
 \begin{figure}
   \begin{mathpar}
-    \infer*[left=V-Int]
+    \infer[S-Int]
+    {
+    l \leq l'
+    }
+    {
+    \textsf{int}^l \leq \textsf{int}^{l'}
+    }
+
+    \infer[S-Abs]
+    {
+    t_{in}' \leq t_{in} \\
+    t_{out} \leq t_{out}' \\
+    l \leq l'
+    }
+    {
+    (t_{in} \to t_{out})^l \leq (t_{in}' \to t_{out}')^{l'}
+    }
+  \end{mathpar}
+  \caption{Rules for subtyping in \langName.}
+  \label{fig:subtype_rules}
+\end{figure}
+
+For the sake of being able to discuss how strong labels are, we follow the
+example of Simonet and Pottier~\cite{InfoFlowML} and define a ``guards''
+relation between labels and types, denoted $l \guards t$, which states that the
+label $l$ is at least as restrictive as the label on the type $t$.  In other
+words, this requires that the type has at least $l$ for it's security label.
+The two rules regarding this relationships for both types in our language are
+given in Figure~\ref{fig:guard_rules}.
+
+\begin{figure}
+  \begin{mathpar}
+    \infer[G-Int]
+    {
+    l \leq l'
+    }
+    {
+    l \guards \textsf{int}^{l'}
+    }
+
+    \infer[G-Int]
+    {
+    l \leq l'
+    }
+    {
+    l \guards (* \to *)^{l'}
+    }
+  \end{mathpar}
+  \caption{Rules for labels guarding types in \langName.}
+  \label{fig:guard_rules}
+\end{figure}
+
+\begin{figure}
+  \begin{mathpar}
+    \infer[V-Int]
     { }
@@
 
-    \infer*[left=V-Var]
+    \infer[V-Var]
     {\Gamma(x) = t}
@@
 
-    \infer*[left=V-Abs]
+    \infer[V-Abs]
     {\typeRule{pc, \Gamma[x \mapsto t'][f \mapsto (t \to t')^l]}{e}{t}}
@@
 
-    \infer*[left=V-Bracket]
+    \infer[V-Bracket]
     {
@@
     l \in H \\
-    l \lhd t
+    l \guards t
     }
@@
 
-    \infer*[left=V-Sub]
+    \infer[V-Sub]
     {
@@
   \caption{Typing judgements for values in \langName.}
+  \label{fig:value_types}
 \end{figure}
 
-The type system for our simple language is given in Figure~\ref{fig:type_rules}.
+\begin{figure}
+  \begin{mathpar}
+    \infer[E-Value]
+    {
+    \typeRule{\Gamma}{v}{t}
+    }
+    {
+    \typeRule{*, \Gamma}{v}{t}
+    }
+
+    \infer[E-App]
+    {
+    \typeRule{\Gamma}{v_1}{(t' \to t)^l} \\
+    \typeRule{\Gamma}{v_2}{t'} \\
+    l \guards t
+    }
+    {
+    \typeRule{pc, \Gamma}{v_1 v_2}{t}
+    }
+
+    \infer[E-If]
+    {
+    \typeRule{pc, \Gamma}{e_{test}}{\textsf{int}^l} \\
+    \typeRule{pc \sqcup l, \Gamma}{e_{true}}{t} \\
+    \typeRule{pc \sqcup l, \Gamma}{e_{false}}{t} \\
+    l \guards t
+    }
+    {
+    \typeRule{pc, \Gamma}{\textsf{if } e_{test} \textsf{ } e_{true} \textsf{ } e_{false}}{t}
+    }
+
+    \infer[E-Let]
+    {
+    \typeRule{\Gamma}{v}{s} \\
+    \typeRule{pc, \Gamma[x \mapsto s]}{e}{t} \\
+    }
+    {
+    \typeRule{pc, \Gamma}{\textsf{let } x := s \textsf{ in } e}{t}
+    }
+
+    \infer[E-Sub]
+    {
+    \typeRule{pc, \Gamma}{e}{t'} \\
+    t' \leq t
+    }
+    {
+    \typeRule{pc, \Gamma}{e}{t}
+    }
+
+    \infer[E-Bracket]
+    {
+    \typeRule{pc \sqcup pc', \Gamma}{e_1}{t} \\
+    \typeRule{pc \sqcup pc', \Gamma}{e_2}{t} \\
+    pc' \in H \\
+    pc' \guards t
+    }
+    {
+    \typeRule{pc, \Gamma}{<e_1 | e_2>}{t}
+    }
+  \end{mathpar}
+  \caption{Typing judgements for expressions in \langName.}
+  \label{fig:expr_types}
+\end{figure}
+
+The type system for our simple language is given in Figure~\ref{fig:value_types}.
 

### 2013-12-18 paper/paper.tex

@@
 \usepackage{amssymb}
+\usepackage{amsthm}
 %\usepackage{float}
@@
 \newcommand{\langName}[0]{TinyML\textsuperscript{2}}
-\newcommand{\typeRule}[3]
-{#1 \textsf{ } \vdash \textsf{ } #2 \textsf{ } \colon \textsf{ } #3}
+\newcommand{\origLang}[0]{CoreML\textsuperscript{2}}
+\newcommand{\typeRule}[3]{#1 \vdash #2 \colon #3}
 \newcommand{\guards}[0]{\lhd}
+\newcommand{\lift}[1]{\lfloor #1 \rfloor}
+
+\theoremstyle{plain}
+\newtheorem{thm}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 
-Citing something for now to make an error message go away~\cite{Pierce:SF}.
 Abstract goes here, summarize what this paper covers and what we ended up with.
@@
 
-\begin{figure}
+\begin{figure}[h!]
   \begin{mathpar}
@@
 
-\begin{figure}
+Figure~\ref{fig:value_types} includes typing judgements for values and
+Figure~\ref{fig:expr_types} includes typing judgements for expressions.
+
+\begin{figure}[h!]
   \begin{mathpar}
@@
 
-The type system for our simple language is given in Figure~\ref{fig:value_types}.
-
 \section{Proofs}
 
-I'm imagining this section could outline the human version of the proofs we'd
-need to do a full noninterference proof.
+Before discussing the task of mechanizing a noninterference theorem for
+\langName, it is useful to first view what the natural language proof would
+look like first.  We first start by proving some supporting lemmas and finish
+with a proof of noninterference for the language.  Since \langName is a stripped
+down version of \origLang, the proof will look like a simplified version of the
+proof found in Pottier and Simonet's paper~\cite{InfoFlowML}.
+
+234101415
+
+\begin{lemma}[Soundness]
+  \label{lem:soundness}
+  Let $i \in \{1, 2\}$. If $e \to e'$, then $\lift{e}_i \to \lift{e'}_i$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}
+  \label{lem:stuck_bracket}
+  If $e$ is stuck, then $\lift{e}_i$is stuck for some $i \in \{1, 2\}$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}[Completeness]
+  \label{lem:stuck_bracket}
+  If $e$ is stuck, then $\lift{e}_i$is stuck for some $i \in \{1, 2\}$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}[Substitution]
+  \label{lem:stuck_bracket}
+  Assume $\typeRule{ }{v}{s}$.  Then, $\typeRule{\Gamma[x \mapsto s]}{v'}{t}$
+  implies $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.  Also $\typeRule{pc,
+  \Gamma[x \mapsto s]}{e}{t}$ implies $\typeRule{pc, \Gamma}{e[x \Leftarrow
+  v]}{t}$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{lemma}
+  Let $H$ be an upward-closed subset of the label lattice $\mathcal{L}$.  Let $l
+  \not\in H$.  If, using $H$ for the set of high labels, $\typeRule{
+  }{e}{\textsf{int}^l}$ and $e \to^* v$ then $\lift{v}_1 = \lift{v}_2$.
+\end{lemma}
+\begin{proof}
+  TODO
+\end{proof}
+
+\begin{thm}[Noninterference]
+  Choose $l, h \in \mathcal{L}$ such that $h \not\leq l$.  Let $h \guards t$.
+  Assume $\typeRule{(x \mapsto t)}{e}{\textsf{int}^l}$, where $e$is a \langName
+  expression.  If, for all $i \in \{1, 2\}$, $\typeRule{ }{v_i}{t}$ and $e[x
+  \Leftarrow v_i] \to^* v_i'$ hold, then $v_1' = v_2'$.
+\end{thm}
+\begin{proof}
+  TODO
+\end{proof}
 

### 2013-12-18 paper/paper.tex
Proofs proofs proofs for the type system.

@@
 \newtheorem{lemma}{Lemma}
+\theoremstyle{definition}
+\newtheorem*{case}{Case}
 
@@
     \infer[V-Var]
-    {\Gamma(x) = t}
+    {t \in \Gamma(x)}
     {\typeRule{\Gamma}{x}{t}}
@@
 with a proof of noninterference for the language.  Since \langName is a stripped
-down version of \origLang, the proof will look like a simplified version of the
-proof found in Pottier and Simonet's paper~\cite{InfoFlowML}.
-
-234101415
+down version of \origLang, the proofs will look like a simplified version of the
+proofs found in Pottier and Simonet's paper~\cite{InfoFlowML}.
 
@@
 
+\begin{lemma}
+  \label{lem:guard_subtype}
+  If $l' \leq l$ and $l \guards t$ and $t \leq t'$, then $l' \guards t'$.
+\end{lemma}
+\begin{proof}
+  $l \guards t$ is equivalent to $l \leq level(t)$ where $level$ is a function
+  that takes a type and returns the label on that type.  It's simple to see that
+  $level$ is covariant in its argument.  The result follows.
+\end{proof}
+
+\begin{lemma}[Weakening]
+  \label{lem:weakening}
+  $pc' \leq pc$ and $\typeRule{pc, \Gamma}{e}{t}$ imply $\typeRule{pc',
+  \Gamma}{e}{t}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\typeRule{pc, \Gamma}{e}{t}$.  By the
+  monotonicity of $\sqcup$, rule \textsc{V-Sub}, Lemma~\ref{lem:guard_subtype},
+  and the inductive hypothesis, it can be checked that the premise remains valid
+  when $pc$ is lowered.
+\end{proof}
+
+\begin{lemma}[Projection]
+  \label{lem:projection}
+  Let $i \in \{1, 2\}$.  If $\typeRule{\Gamma}{v}{t}$ then
+  $\typeRule{\Gamma}{\lift{v}_i}{t}$.  If $\typeRule{pc, \Gamma}{e}{t}$ then
+  $\typeRule{pc, \Gamma}{\lift{e}_i}{t}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation for $v$ and $e$.  The only interesting case is
+  that of \textsc{E-Bracket}, since other cases end up either being trivial or
+  using the inductive hypothesis for a case of \textsc{E-Bracket}.  The
+  expression is $<e_1 | e_2>$.  Since $pc \sqcup pc' \leq pc$, by
+  Lemma~\ref{lem:weakening} and the inductive hypothesis, we can prove the
+  typing judgements for both branches.
+\end{proof}
+
+\begin{lemma}[Guard]
+  \label{lem:guard}
+  If $\typeRule{\Gamma}{< v_1 | v_2 >}{t}$ then there exists $pc' \in H$ s.t.
+  $pc' \guards t$.
+\end{lemma}
+\begin{proof}
+  Through Lemma~\ref{lem:guard_subtype}, we can see that we can always construct
+  a typing derivation that does not end with \textsc{V-Sub}.  As a result, the
+  derivation can always end with \textsc{V-Bracket}, which has the premises $pc'
+  \in H$ and $pc' \guards t$.
+\end{proof}
+
 \begin{lemma}[Substitution]
@@
 \begin{proof}
-  TODO
+  We can prove both statements by induction on the typing derivation for $v'$
+  and $e$.  In this proof, we focus only on a subset of the cases.
+
+  \begin{case}[\textsc{V-Int}]
+    This is trivial, since $x$ does not get substituted in $v'$ and so the
+    statement becomes a tautology.
+  \end{case}
+
+  \begin{case}[\textsc{V-Var}]
+    If $v'$ is $x$, then we are in the case where $s \in t$.  Thus, we can
+    conclude that given $\typeRule{ }{v}{s}$, $\typeRule{\Gamma[x \mapsto
+    s]}{v'}{t}$ by \textsc{V-Var}.  Since $v'[x \Leftarrow v] = v$, we can
+    conclude $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.
+
+    Otherwise, we are in the case where $v'$ is some variable $y \neq x$.
+    Then, given that $v'[x \Leftarrow v] = v' = y$ and since $\Gamma(y) =
+    \Gamma[x \mapsto s]$, we can see that $\typeRule{\Gamma[x \mapsto
+    s]}{v'}{t}$ implies that $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.
+  \end{case}
+
+  \begin{case}[\textsc{V-Abs}]
+    The premise in this case becomes $\typeRule{pc, \Gamma[x \mapsto s][y
+    \mapsto t'][f \mapsto (t' \to t'')^l]}{e'}{t''}$.
+
+    Let us assume that $x$, $f$, and $y$ are distinct, since otherwise this
+    would reduce to a similar situation to that of \textsc{V-Var} when the
+    variable substitution leaves $v'$ unchanged, which is trivial.  Since they
+    are distinct, $\Gamma[x \mapsto s][y \mapsto t'][f \mapsto (t' \to t'')^l]$
+    is equivalent to $\Gamma[y \mapsto t'][f \mapsto (t' \to t'')^l][x \mapsto
+    s]$ and so we can apply the inductive hypothesis and \textsc{V-Abs} to show
+    $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.
+  \end{case}
+
+  \begin{case}[\textsc{V-Bracket}]
+    In this case the premises are of the form $\typeRule{\Gamma[x \mapsto
+    s]}{v'_i}{t}$.  By Lemma~\ref{lem:projection}, we know that
+    $\typeRule{\Gamma}{v}{s}$ implies $\typeRule{\Gamma}{\lift{v}_i}{s}$.  By
+    the inductive hypothesis, we can then say that $\typeRule{\Gamma}{v'_i[x
+    \Leftarrow v]}{t}$ holds.  Thus, we can apply \textsc{V-Bracket} to obtain
+    $\typeRule{\Gamma}{<v'_1[x \Leftarrow v] | v'_2[x \Leftarrow v]>}{t}$,
+    proving the statement.
+  \end{case}
+
+  \begin{case}[\textsc{V-Sub}, \textsc{E-Value}, \textsc{E-App}, \textsc{E-If},
+    \textsc{E-Let}, \textsc{E-Sub}]
+    Applying the inductive hypothesis on the premises and applying the
+    associated rule proves these cases.
+  \end{case}
+
+  \begin{case}[\textsc{E-Bracket}]
+    This case is similar to the case of \textsc{V-Bracket}.
+    Lemma~\ref{lem:guard} allows us to apply \textsc{E-Bracket} where necessary.
+  \end{case}
 \end{proof}
@@
 
-We would like to thank Jon Sailor from Facebook for helping us over the phone
-work through some of the more difficult problems we ran into using Coq.
+We would like to thank Jon Sailor for chatting with us on the phone to help work
+through some of the more difficult problems we ran into using Coq.
 

### 2013-12-18 paper/paper.tex
Finally figured out how to format operational semantics rules...
Operational semantics written.
Fixes to semantics and proof of first 2 lemmas.

@@
 \usepackage{mathpartir}
+\usepackage{amsmath}
 \usepackage{amssymb}
@@
 any interesting similarities or differences between us and Pottier and Simonet.
+
+\begin{figure}
+  Basic reductions
+  \begin{align}
+    (\textsf{fix} f . \lambda x . e) v
+    \textsf{ } \to \textsf{ } &
+    e[x \Leftarrow v][f \Leftarrow \textsf{fix} f . \lambda x . e]
+    \tag{$\beta$}
+    \label{eq:beta}
+    \\
+    \textsf{if } 1 \textsf{ } e_1 \textsf { } e_2
+    \textsf{ } \to \textsf{ } &
+    e_1
+    \tag{if-true}
+    \label{eq:if-true}
+    \\
+    \textsf{if } (k \neq 1) \textsf { } e_1 \textsf { } e_2
+    \textsf{ } \to \textsf{ } &
+    e_2
+    \tag{if-false}
+    \label{eq:if-false}
+    \\
+    \textsf{let }  x := v \textsf{ in } e
+    \textsf{ } \to \textsf{ } &
+    e[x \Leftarrow v]
+    \tag{let}
+    \label{eq:let}
+  \end{align}
+
+  Lifting
+  \begin{align}
+    < v_1 | v_2 > v
+    \textsf{ } \to \textsf{ } &
+    < v_1 \lift{v}_1 | v_2 \lift{v}_2 >
+    \tag{lift-app}
+    \label{eq:lift-app}
+    \\
+    \textsf{if } < v_1 | v_2 > \textsf{ } e_1 \textsf{ } e_2
+    \textsf{ } \to \textsf{ } &
+    < \textsf{if } v_1 \textsf{ } \lift{e_1}_1 \textsf{ } \lift{e_2}_1 |
+    \textsf{ if } v_2 \textsf{ } \lift{e_1}_2 \textsf{ } \lift{e_2}_2 >
+    \tag{lift-if}
+    \label{eq:lift-if}
+    \\
+    &
+    \infer*
+    {
+    e_1 \to e_1' \\
+    }
+    {
+    <e_1 | e_2> \textsf{ } \to \textsf{ } <e_1' | e_2>
+    }
+    \tag{bracket-left}
+    \label{eq:bracket-left}
+    \\
+    &
+    \infer*
+    {
+    e_2 \to e_2' \\
+    }
+    {
+    <e_1 | e_2> \textsf{ } \to \textsf{ } <e_1 | e_2'>
+    }
+    \tag{bracket-right}
+    \label{eq:bracket-right}
+  \end{align}
+  \caption{Operational semantics for \langName.}
+  \label{fig:operational_semantics}
+\end{figure}
 
@@
   \label{lem:soundness}
-  Let $i \in \{1, 2\}$. If $e \to e'$, then $\lift{e}_i \to \lift{e'}_i$.
+  Let $i \in \{1, 2\}$. If $e \to e'$, then $\lift{e}_i \to^*
+  \lift{e'}_i$.\footnote{Notice that this is not the same statement as the
+  Soundness lemma in Simonet and Pottier's paper~\cite{InfoFlowML}, which does
+  not give a variable number of steps for the conclusion.  We will discuss this
+  change in Section~\ref{sec:mechanizing}.}
 \end{lemma}
 \begin{proof}
-  TODO
+  This can be seen by examining each of the operational semantics rules in
+  Figure~\ref{fig:operational_semantics}.
 \end{proof}
@@
 \begin{proof}
-  TODO
+  By induction on the structure of $e$.
+
+  \begin{case}[$e = v$]
+    In this case, $e$ is not stuck.
+  \end{case}
+
+  \begin{case}[$e = v_1 v_2$]
+    If $e$ is stuck and is an application, this means that neither
+    \ref{eq:beta} nor \ref{eq:lift-app} applies.  This means that $v_1$ can
+    neither be a fixpoint function or some bracketed value.  As a result,
+    $\lift{v_1}_i$ can not be a fixpoint function, and so it is stuck.
+  \end{case}
+
+  \begin{case}[$e = \textsf{ if } v \textsf{ } e_1 \textsf{ } e_2$ or $e =
+    \textsf{ let } x := v \textsf{ in } e$]
+    These are similar to the previous case.  The value $v$ can not be an integer
+    and so the lifted version of $v$ can not be an integer, meaning the lifted
+    expression is also stuck.
+  \end{case}
+
+  \begin{case}[$e = < e_1 | e_2 >$]
+    Since $e$ is stuck, then by \ref{eq:bracket-left} and \ref{eq:bracket-right}
+    we know that both $e_1$ and $e_2$ must be stuck.  Since $e$ is not a final
+    value, we know either $e_1$ or $e_2$ must not be a final value and, as a
+    result, stuck.
+  \end{case}
 \end{proof}
@@
 \begin{lemma}[Completeness]
-  \label{lem:stuck_bracket}
-  If $e$ is stuck, then $\lift{e}_i$is stuck for some $i \in \{1, 2\}$.
+  \label{lem:completeness}
+  Assume $\lift{e}_i \to^* a_i$ for all $i \in \{1, 2\}$.  Then, there exists a
+  final value $a$ such that $e \to^* a$.
 \end{lemma}
@@
 \section{Mechanizing Proofs}
+\label{sec:mechanizing}
 

### 2013-12-18 paper/paper.tex
Added syntax and rearranged structure a bit.  Also figured out a more terse way to make a space in a math environment.

@@
 
-\section{Operational Semantics}
+\section{\langName}
+
+\begin{figure}
+  \begin{align}
+    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \\
+    a & ::= v \tag{answers} \\
+    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e_1 \: e_2 \: | \: \textsf{let
+    } x := v \textsf{ in } e \tag{expressions} \\
+  \end{align}
+  \caption{Syntax of \langName}
+  \label{fig:syntax}
+\end{figure}
+
+\subsection{Operational Semantics}
 
@@
     (\textsf{fix} f . \lambda x . e) v
-    \textsf{ } \to \textsf{ } &
+    \: \to \: &
     e[x \Leftarrow v][f \Leftarrow \textsf{fix} f . \lambda x . e]
@@
     \\
-    \textsf{if } 1 \textsf{ } e_1 \textsf { } e_2
-    \textsf{ } \to \textsf{ } &
+    \textsf{if } 1 \: e_1 \textsf { } e_2
+    \: \to \: &
     e_1
@@
     \textsf{if } (k \neq 1) \textsf { } e_1 \textsf { } e_2
-    \textsf{ } \to \textsf{ } &
+    \: \to \: &
     e_2
@@
     \textsf{let }  x := v \textsf{ in } e
-    \textsf{ } \to \textsf{ } &
+    \: \to \: &
     e[x \Leftarrow v]
@@
     < v_1 | v_2 > v
-    \textsf{ } \to \textsf{ } &
+    \: \to \: &
     < v_1 \lift{v}_1 | v_2 \lift{v}_2 >
@@
     \\
-    \textsf{if } < v_1 | v_2 > \textsf{ } e_1 \textsf{ } e_2
-    \textsf{ } \to \textsf{ } &
-    < \textsf{if } v_1 \textsf{ } \lift{e_1}_1 \textsf{ } \lift{e_2}_1 |
-    \textsf{ if } v_2 \textsf{ } \lift{e_1}_2 \textsf{ } \lift{e_2}_2 >
+    \textsf{if } < v_1 | v_2 > \: e_1 \: e_2
+    \: \to \: &
+    < \textsf{if } v_1 \: \lift{e_1}_1 \: \lift{e_2}_1 \: |
+    \textsf{ if } v_2 \: \lift{e_1}_2 \: \lift{e_2}_2 >
     \tag{lift-if}
@@
     {
-    <e_1 | e_2> \textsf{ } \to \textsf{ } <e_1' | e_2>
+    <e_1 | e_2> \: \to \: <e_1' | e_2>
     }
@@
     {
-    <e_1 | e_2> \textsf{ } \to \textsf{ } <e_1 | e_2'>
+    <e_1 | e_2> \: \to \: <e_1 | e_2'>
     }
@@
 
-\section{Type System}
+\subsection{Type System}
 
@@
     {
-    \typeRule{pc, \Gamma}{\textsf{if } e_{test} \textsf{ } e_{true} \textsf{ } e_{false}}{t}
+    \typeRule{pc, \Gamma}{\textsf{if } e_{test} \: e_{true} \: e_{false}}{t}
     }
@@
 
-\section{Proofs}
+\section{Proof of Noninterference}
 
-Before discussing the task of mechanizing a noninterference theorem for
-\langName, it is useful to first view what the natural language proof would
-look like first.  We first start by proving some supporting lemmas and finish
-with a proof of noninterference for the language.  Since \langName is a stripped
-down version of \origLang, the proofs will look like a simplified version of the
-proofs found in Pottier and Simonet's paper~\cite{InfoFlowML}.
+Before discussing the task of mechanizing a noninterference proof for \langName,
+it is useful to first view what the natural language proof would look like
+first.  We first start by proving some supporting lemmas about the language's
+semantics and type system, and finish with a proof of noninterference for the
+language.  Since \langName is a stripped down version of \origLang, the proofs
+will look like a simplified version of the proofs found in Pottier and Simonet's
+paper~\cite{InfoFlowML}.
 
@@
 
-  \begin{case}[$e = \textsf{ if } v \textsf{ } e_1 \textsf{ } e_2$ or $e =
+  \begin{case}[$e = \textsf{ if } v \: e_1 \: e_2$ or $e =
     \textsf{ let } x := v \textsf{ in } e$]
@@
     Since $e$ is stuck, then by \ref{eq:bracket-left} and \ref{eq:bracket-right}
-    we know that both $e_1$ and $e_2$ must be stuck.  Since $e$ is not a final
-    value, we know either $e_1$ or $e_2$ must not be a final value and, as a
+    we know that both $e_1$ and $e_2$ must be stuck.  Since $e$ is not an
+    answer, we know either $e_1$ or $e_2$ must not be an answer and, as a
     result, stuck.
@@
   \label{lem:completeness}
-  Assume $\lift{e}_i \to^* a_i$ for all $i \in \{1, 2\}$.  Then, there exists a
-  final value $a$ such that $e \to^* a$.
+  Assume $\lift{e}_i \to^* a_i$ for all $i \in \{1, 2\}$.  Then, there exists an
+  answer $a$ such that $e \to^* a$.
 \end{lemma}

### 2013-12-18 paper/paper.tex
Very short and sweet abstract.

@@
 
-Abstract goes here, summarize what this paper covers and what we ended up with.
+In this paper we discuss our experience trying to mechanize a proof of
+noninterference for a simple language with security labels on the types.
+Although we did not complete the full proof because we ran out of time, we were
+able to find small inaccuracies and subtleties in the proofs from the paper we
+based our proof on.  We discuss what would be involved in completing the proof
+and how this relates to other work.
 
@@
 
-\begin{figure}
+Our toy language for proving noninterference, \langName, is a simplified version
+of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
+limited amount of time to work on this project, we limited ourselves to a subset
+of the language that drops the following features:
+
+\begin{itemize}
+  \item References and associated memory operations.
+  \item Paired values.
+  \item Exceptions.
+  \item Sequencing operations.
+\end{itemize}
+
+While the language becomes much much less interesting for actual programming, we
+feel that these features do not fundamentally change the problem of proving
+noninterference.  In order to mechanize a proof of noninterference for a
+language closer to \origLang, it would require simply adding mechanized proofs
+for the additional cases where appropriate.
+
+\textbf{TODO: Talk about what the bracketed expressions and values are.}
+
+Furthermore, primarily due to a matter of taste, we decided to replace the
+\textsf{case} construct from \origLang for an \textsf{if} construct, which
+evaluates the first expression if the test evaluates to 1 and evaluates to the
+second expression otherwise.
+
+Through these changes, our language has the syntax shown in
+Figure~\ref{fig:syntax}.  As shown, final states, answers, in the language are
+now simple values because there are no exceptions.
+
+\begin{figure}[h!]
   \begin{align}
-    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \\
-    a & ::= v \tag{answers} \\
-    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e_1 \: e_2 \: | \: \textsf{let
-    } x := v \textsf{ in } e \tag{expressions} \\
+    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \: |
+    \: <v | v> \\
+    a & ::= v \tag{answers} \: | \: <a | a> \\
+    e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
+    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e> \\
   \end{align}
@@
 
+In the following two subsections, we introduce the operational semantics and
+type system for the language.
+
 \subsection{Operational Semantics}
 
-Outline the operational semantics of the baby language we developed.  Outline
-any interesting similarities or differences between us and Pottier and Simonet.
+\langName has an operational semantics that should be familiar to the reader who
+has seen a single step semantics for a basic lambda calculus.  The interesting
+additions are the same as those in \origLang, where there are additional rules
+for handling the bracketed values.
 
-\begin{figure}
+\textbf{TODO: Talk about what the rules do at a high level.}
+
+\begin{figure}[h!]
   Basic reductions
@@
 
+\textbf{TODO: Talk more about the rules and what they're doing.  Specifically,
+discuss what the heck H and pc are doing.}
+
 \section{Proof of Noninterference}
@@
 paper~\cite{InfoFlowML}.
+
+\textbf{TODO: Add in text between lemmas and theorems discussing what each one
+is at a high level.}
 
@@
 \begin{proof}
-  TODO
+  \textbf{TODO}
 \end{proof}

### 2013-12-18 paper/paper.tex
Added a reference to nuprl.  Wrote a basic introduction.

@@
 In this paper we discuss our experience trying to mechanize a proof of
-noninterference for a simple language with security labels on the types.
-Although we did not complete the full proof because we ran out of time, we were
-able to find small inaccuracies and subtleties in the proofs from the paper we
-based our proof on.  We discuss what would be involved in completing the proof
-and how this relates to other work.
+noninterference for a simple language with security labels, \langName.  Although
+we did not complete the full noninterference proof because we ran out of time,
+we were able to find small inaccuracies and subtleties in the proofs from the
+paper we based our proof on.  We discuss what would be involved in completing
+the proof and how this relates to other work.
 
@@
 
-Introduce idea and outline what it will require us to do.  Explain where we're
-starting from (Pottier and Simonet) at a high level.  Justify why this is an
-interesting problem that needs tackling.
+Verified systems have been growing in popularity recently with the rise of proof
+assistants like Coq and NuPRL~\cite{coqart,nuprl}.  These systems allow a
+programmer to write programs and have proofs, which are checked and compiled by
+the system, which can assert various properties about the programs.  As we will
+discuss in Section~\ref{sec:related}, these systems have been used to formalize,
+prove correct, and implement systems that have verified properties.  This allows
+for programmers to be more confident in their languages and programs, ensuring
+that the properties they want are actually properties of their work.
+
+Interestingly, as far as we know there has not been work to formalize and verify
+a language, like Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow},
+that has a type system for enforcing information flow security policies.  For
+languages like these and other systems related to security, it is desirable for
+the program to be accompanied with a proof of the security properties it
+enforces.  Our goal is to make progress towards verifying the correctness of a
+compiler for a language with information flow types that programmers will want
+to use.
+
+A verified compiler for a language like Jif would require a proof of
+noninterference, which asserts that any program in the language will never
+produce different outputs for low security items if only the high security input
+values are varied.  Thus, in this paper we discuss our experience working
+towards proving this theorem for a simple information flow type system as a
+first step towards creating a verified information flow compiler.
+
+Our language, \langName,  and proof are a simplified version of what is
+presented in Simonet and Pottier's work on information flow for an ML-like
+language, \origLang~\cite{InfoFlowML}.  This proof proceeds by designing a
+language that includes values which represent two different alternatives,
+representing the different values of high security items in two different
+executions of the program.  The proof then becomes a matter of proving that if
+the program is typed with a low security label, this does not produce a value
+that could be different between two different executions of the program.
+
+In this paper we will present our simplified language, \langName,
+along with the semantics and type system (Section~\ref{sec:lang}).  We will then
+present a natural language, human readable, proof of noninterference for the
+language and discuss our experiences with mechanizing this proof in the Coq
+proof assistant (Sections~\ref{sec:proof} and \ref{sec:mechanizing}).  Finally,
+we will discuss related work, next steps, and conclude
+(Sections~\ref{sec:related}, \ref{sec:future}, and \ref{sec:conclusion}).
 
 \section{\langName}
+\label{sec:lang}
 
@@
 \section{Proof of Noninterference}
+\label{sec:proof}
 
@@
 \section{Related Work}
+\label{sec:related}
 
@@
 \section{Future Work}
+\label{sec:future}
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2013-12-18 paper/paper.tex
Added a slightly more informative plan for related work and acknowledged the authors of SF for writing an awesome book.

@@
 
+Discuss:
+
+\begin{itemize}
+  \item Proof carrying code
+  \item Proof assistants
+  \item Verified languages and compilers
+  \item Information flow languages
+\end{itemize}
+
 \section{Future Work}
@@
 
-We would like to thank Jon Sailor for chatting with us on the phone to help work
-through some of the more difficult problems we ran into using Coq.
+We would like to thank Jon Sailor for chatting with us on the phone for hours to
+help work through some of the more difficult problems we ran into using Coq.
+Also, we would like to thank the authors of Software
+Foundations~\cite{Pierce:sf}, which was the primary resource for learning Coq
+for both authors.
 

### 2013-12-18 paper/paper.tex
A bit of fleshing out of the operational semantics section and small tweaks elsewhere.

@@
 In this paper we discuss our experience trying to mechanize a proof of
-noninterference for a simple language with security labels, \langName.  Although
-we did not complete the full noninterference proof because we ran out of time,
-we were able to find small inaccuracies and subtleties in the proofs from the
-paper we based our proof on.  We discuss what would be involved in completing
-the proof and how this relates to other work.
+noninterference for a simple language with security labels, \langName.  This
+marks the first step towards writing a verified compiler for a language with
+information flow types.  Although we did not complete the full noninterference
+proof because we ran out of time, we were able to find small inaccuracies and
+subtleties in the proofs from the paper we based our proof on.  We discuss what
+would be involved in completing the proof and how this relates to other work.
 
@@
 
-\textbf{TODO: Talk about what the bracketed expressions and values are.}
-
 Furthermore, primarily due to a matter of taste, we decided to replace the
@@
 second expression otherwise.
+
+\langName, like \origLang, includes angle bracketed expressions, which represent
+values or expressions which could differ between two runs of the program.  As we
+will see when discussing the type system, this corresponds to values and
+expressions that have high security labels on their types.
 
@@
     e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
-    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e> \\
+    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e>
   \end{align}
@@
 additions are the same as those in \origLang, where there are additional rules
-for handling the bracketed values.
-
-\textbf{TODO: Talk about what the rules do at a high level.}
+for handling the bracketed values.  The reduction rules for the semantics are
+given in Figure~\ref{fig:operational_semantics}.
 
@@
 \end{figure}
+
+At a high level, these bracketed expressions are evaluated by evaluating one of
+the branches, as shown in rules \ref{eq:bracket-left} and
+\ref{eq:bracket-right}, or by expanding the brackets around the parent
+expression, passing the appropriate branch of the bracketed subvalue to each
+branch of the resulting expression, as shown in rules \ref{eq:lift-app} and
+\ref{eq:lift-if}.  In these rules, we use the notation $\lift{x}_i$ to denote
+the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value.
+
+We prove soundness and completeness lemmas for these rules in
+Section~\ref{sec:proof}.
 

### 2013-12-18 paper/paper.tex
Added supporting text to both the type system and the proofs.  Still need to give a basic outline of types in the type system section.

@@
 \ref{eq:lift-if}.  In these rules, we use the notation $\lift{x}_i$ to denote
-the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value.
+the left ($i = 1$) or right ($i = 2$) branch of a bracketed expression or value
+$x$.
 
@@
 
+\textbf{TODO: Give definition of types in the language and discuss stuff that
+was dropped from \origLang}
+
 The language's type system includes subtyping, which allows data that have a
-lower label to flow into higher labelled data.  The rules for subtyping are
+lower label to flow into higher labelled data.  We use the notatoin $x \leq y$
+to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
+types (as long as both are the same kind of value).  The rules for subtyping are
 given in Figure~\ref{fig:subtype_rules}.  Rule \textsc{S-Int} states that
@@
 
-For the sake of being able to discuss how strong labels are, we follow the
-example of Simonet and Pottier~\cite{InfoFlowML} and define a ``guards''
-relation between labels and types, denoted $l \guards t$, which states that the
-label $l$ is at least as restrictive as the label on the type $t$.  In other
-words, this requires that the type has at least $l$ for it's security label.
-The two rules regarding this relationships for both types in our language are
-given in Figure~\ref{fig:guard_rules}.
+For the sake of being able to discuss how strong labels are in relation to a
+type, we follow the example of Simonet and Pottier~\cite{InfoFlowML} and define
+a ``guards'' relation between labels and types, denoted $l \guards t$, which
+states that the label $l$ is at least as restrictive as the label on the type
+$t$.  In other words, this requires that the type has at least $l$ for it's
+security label.  The two rules regarding this relationships for both types in
+our language are given in Figure~\ref{fig:guard_rules}.
 
@@
 
-Figure~\ref{fig:value_types} includes typing judgements for values and
-Figure~\ref{fig:expr_types} includes typing judgements for expressions.
+Figure~\ref{fig:value_types} includes typing judgements for values in the
+language, which are mostly straight forward with the exception of
+\textsc{V-Bracket}, which we discuss below along with the type rules for
+expressions.
 
@@
 \end{figure}
+
+The typing rules for expressions in the language are given in
+Figure~\ref{fig:expr_types}.  Unlike a typical typing judgment, we include an
+additional piece of context information, the $pc$, which is a label that
+describes what level of label can influence the control structure decisions to
+reach the current expression in the program.  This is a standard extension for
+languages with information flow types, like Jif.
+
+The interesting rules in the type system are \textsc{V-Bracket},
+\textsc{E-Bracket}, and \textsc{E-If}.  \textsc{E-If} is a standard rule for
+information flow type systems, stating that the type determinations for each
+branch must have a $pc$ equivalent to the security of the original $pc$ along
+with the security of the label on the value in the test, $l$.
+
+\textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
+from the set $H$ which representes an upward closed sublattice of the labels in
+the system representing what is considered to be ``high'' security for the
+purposes of the proof of noninterference, discussed later.  These rules require
+that the types of each branch of a bracketed value or expression can be typed
+with a high security label.  Furthermore, \textsc{E-Bracket} requires that the
+$pc$ of the subexpressions must be at least as high as this same label,
+requiring that we treat everything in the subexpression as affected by high
+security information.
 
@@
 
-\textbf{TODO: Talk more about the rules and what they're doing.  Specifically,
-discuss what the heck H and pc are doing.}
-
 \section{Proof of Noninterference}
@@
 
-\textbf{TODO: Add in text between lemmas and theorems discussing what each one
-is at a high level.}
+We start with soundness and completeness lemmas for the operational semantics,
+showing that this language can accurately describe the process of evaluating two
+different runs of a language where bracketed expressions are high security
+values.  Specifically, the image of a valid reduction of an expression is also a
+valid reduction and if both runs of a program evaluate to an answer, then so can
+the bracketed expression in \langName.
 
@@
 
+Next, we prove lemmas related to the type system.  We start with a lemma showing
+that $\guards$ interacts well with subtyping and a weakening lemma, showing that
+anything typed at a given $pc$ can be typed at a lower $pc'$ as well.
+
 \begin{lemma}
@@
 
+We now prove a projection lemma, showing that if a top level expression or value
+is well typed, so must either projection of that expression or value to one
+program run or another.
+
 \begin{lemma}[Projection]
@@
 
+Below is a lemma showing that if a bracketed expression is well typed, then
+there must exist a high security label to use for the types and $pc$ of both
+branches.
+
 \begin{lemma}[Guard]
@@
 \end{proof}
+
+Here we prove a standard substitution lemma for the type system.
 
@@
 
+Finally, we prove one final lemma and the noninterference theorem.  The lemma
+shows that if an expression is typed with a low security label, then both
+projections (so the value in either run of the program) are equivalent.
+
 \begin{lemma}
@@
 \end{proof}
+
+Finally, the noninteference theorem shows that if we have an expression that is
+typed with a low label, then replacing any bracketed value in the expression
+with either projection will result in the same value for the overall expression.
+In other words, if we vary any high security item in a low security expression,
+the value must not be affected.
 

### 2013-12-18 paper/paper.tex
Added basic type definitions and addressed changes to function types.

@@
 abstractions, which include fixpoints by default, along with appropriate
-security labels for both.  In \langName, security labels specify confidentiality
-policies, specifying what other data the labelled data can affect.\footnote{It is
-not difficult to imagine how to extend our system to include integrity labels.
-However, for the sake of simplicity, we decided to concern ourselves with only
-confidentiality at first}  Data can only be affected by items at the same label
-or lower (so high data can not affect low labelled data).
+security labels for both, as shown in Figure~\ref{fig:types}.  In \langName,
+security labels specify confidentiality policies, specifying what other data the
+labelled data can affect.\footnote{It is not difficult to imagine how to extend
+our system to include integrity labels.  However, for the sake of simplicity, we
+decided to concern ourselves with only confidentiality at first}  Data can only
+be affected by items at the same label or lower (so high data can not affect low
+labelled data).
 
-\textbf{TODO: Give definition of types in the language and discuss stuff that
-was dropped from \origLang}
+\begin{figure}[h!]
+  \begin{align*}
+    t ::= int^l \: | \: (t -> t)^l
+  \end{align*}
+  \caption{Types for \langName}
+  \label{fig:types}
+\end{figure}
+
+A reader familiar with the type system for \origLang will notice that our type
+system does not include types above the arrow in the function type, for giving
+security bounds on side effects of the function, and the type does not include
+effect labels for exceptions.  These are dropped as a result of the
+simplifications we made in defining \langName, since there are no side effects
+or exceptions in the language.
 

### 2013-12-18 paper/paper.tex
Corrections, more writing, a todo macro, and some explanation of related work and our experience with mechanization.

@@
 \newcommand{\lift}[1]{\lfloor #1 \rfloor}
+\newcommand{\todo}[1]{\textbf{TODO: #1}}
 
@@
   \begin{align}
-    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag{values} \: |
-    \: <v | v> \\
-    a & ::= v \tag{answers} \: | \: <a | a> \\
+    v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag*{ } \\
+    v' & ::= v \: | \: <v | v> \tag{values}\\
+    a & ::= v \tag*{ } \\
+    a' & ::= a \: | \: <a | a> \tag{answers} \\
     e & ::= a \: | \: v v \: | \: \textsf{if } v \: e \: e \: | \: \textsf{let
-    } x := v \textsf{ in } e \tag{expressions} \: | \: <e | e>
+    } x := v \textsf{ in } e \tag*{ } \\
+    e' & ::= e \: | \: <e | e> \tag{expressions}
   \end{align}
-  \caption{Syntax of \langName}
+  \caption{Syntax of \langName.  The prime variants are the full class of
+  expressions, values, and answers.  This construction avoids allowing nested
+  branching values.}
   \label{fig:syntax}
@@
 \begin{proof}
-  \textbf{TODO}
+  First, we establish that if $\lift{e}_i$ reduces to an answer in a finite
+  number of steps, then $e$ does not have an infinite reduction sequence.  The
+  only way this could occur is if $e$ was infinitely reduced over lift rules,
+  because otherwise, since the only reductions on $e$ that correspond to 0
+  reductions on either projection are lift reduction rules.  But as we can see
+  from inspecting the lift reduction rules, we cannot continuously apply these
+  rules exclusively.  Thus, we can conclude that $e$ does not have an infinite
+  series of reduction steps that can be applied to it.
+
+  As a result, we know that $e$ reduces to an answer, because otherwise it would
+  be stuck and this would mean that one of the projections is also stuck by
+  Lemma~\ref{lem:bracket_stuck}, a contradicton.  Thus, we know that if both
+  projections step to an answer, so must the original expression.
 \end{proof}
@@
 
+The following lemma asserts that if a value is well typed, the expression
+containing just that value is also well typed.
+
+\begin{lemma}[Value]
+  \label{lem:value}
+  $\typeRule{pc}{v}{t}$ implies $\typeRule{ }{v}{t}$
+\end{lemma}
+\begin{proof}
+  By induction on the proof of $\typeRule{pc}{v}{t}$.  The cases are either that
+  we have \textsc{E-Value}, where we're done, \textsc{E-Sub}, in which case we
+  apply the inductive hypothesis and the rule \textsc{V-Sub}, or we have
+  \textsc{E-Bracket}, which is proved by the inductive hypothesis and
+  \textsc{V-Bracket}.
+\end{proof}
+
+We also prove a subject reduction lemma, allowing us to assert that if an
+expression is well typed and there is a reduction that can apply to it, then the
+reduced form is also well typed, with the same type.
+
+\begin{lemma}[Subject Reduction]
+  \label{lem:subj_red}
+  Let $e \to e'$.  Assume $\typeRule{pc}{e}{t}$ and $pc \in H$ for all $i \in
+  \{1, 2\}$.  Then $\typeRule{pc}{e'}{t}$.
+\end{lemma}
+\begin{proof}
+  \todo{}
+\end{proof}
+
 Finally, we prove one final lemma and the noninterference theorem.  The lemma
@@
 \begin{lemma}
+  \label{lem:final_support}
   Let $H$ be an upward-closed subset of the label lattice $\mathcal{L}$.  Let $l
@@
 \begin{proof}
-  \textbf{TODO}
+  By Lemma~\ref{lem:subj_red} and \ref{lem:value}, we know that $\typeRule{
+  }{v}{\textsf{int}^l}$ holds.  Since the type is $\textsf{int}^l$, we know that
+  $v$ is either $k$ or $< k_1 | k_2 >$.  However, we know that if it were the
+  latter, then by \textsc{V-Bracket} this would imply there is an $l' \leq l$
+  such that $l' \in H$, a contradiction.  Thus, $v$ must be of the form $k$
+  which is the same under either projection.
 \end{proof}
@@
 \begin{proof}
-  \textbf{TODO}
+  Let $H$ be the upward closure of $\{h\}$ and $v = < v_1 | v_2 >$, since other
+  $v$ would be immediately uninteresting.  By \textsc{V-Bracket}, we know that
+  $\typeRule{ }{v}{t}$ holds given the assumptions.
+  By Lemma~\ref{lem:substitution}, we have $\typeRule{ }{e[x \Leftarrow
+  v]}{\textsf{int}^l}$.  By Lemma~\ref{lem:final_support}, we know that
+  $\lift{e[x \Leftarrow v]}_i = e[x \Leftarrow \lift{v}_i]$, which under the
+  current assumptions reduces to $v_i$.  By Lemma~\ref{lem:completeness} we know
+  that there is an answer $a$ such that $e[x \Leftarrow v] ->^* a$.  By
+  Lemma~\ref{lem:soundness}, $\lift{a}_i = v_i'$ for all $i$ which means that
+  $a$ is a value.  Then, we know by Lemma~\ref{lem:final_support} that $v_1' =
+  v_2'$.
 \end{proof}
@@
 
-Talk about what did/should have been mechanized.  Talk about troubles and/or
-interesting bits here.  Might point out here that we're fairly certain that we
-found an inaccuracy in the original paper.
+As mentioned above, we unfortunately were unable to complete our noninterference
+proof in Coq by the time of this report.  Given more time, we are confident that
+this could be completed however.  So far, we have proven the first few lemmas
+described in Section~\ref{sec:proof}.  This was much more difficult than the
+short natural language proofs would suggest, due to a combination of the
+stricter format required by Coq and the number of items that are elided in the
+proofs because they were considered trivial or obvious.
+
+We found that the process of mechanizing the lemmas that we proved was a useful
+exercise in checking our understanding of our language, statements, and the work
+of Simonet and Pottier which we based our approach off of.  We found ourselves
+often surprised by the number of items that were elided by the proofs in the
+paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
+natural languae but turned out to be 40 lines of Coq code, not including all of
+the supporting definitions and rules.
+
+Our proof of Lemma~\ref{lem:soundness} actually turned out to be instructive
+because it forced us to recognize that the original formulation in Simonet and
+Pottier's paper on \origLang~\cite{InfoFlowML} wasn't quite correct.
+Specifically, it is entirely possible that a step in the operational semantics
+does not translate to a step in the operational semantics under either
+projection, as in the case of the \ref{eq:lift-app} rule, where $\lift{e}_i$ and
+$\lift{e'}_i$ are identical.  Luckily, we are fairly confident, given that we
+were still able to prove Lemma~\ref{lem:bracket_stuck} despite this change, that
+this is not a critical error in the original paper.
+
+\todo{More.}
 
@@
 
-Discuss various references related to our project.
+This project is related to a variety of previous work, specifically in the realm
+of security and verified proofs and programs.  We are following in the footsteps
+of verified languages and compilers, which focus on writing programs and
+language specs that are verified to have the properties the authors want by
+providing the proofs of these properties on the
+definitions~\cite{strecker2002formal,leinenbach2005towards,dave2003compiler,leroy2009formally,leroy2006formal,leroy2009formal,leroy2012compcert,chlipala2010verified,berghofer2004extracting}.
+This has been a popular area, which is primarily focused on more mainstream
+languages and compilers (such as C and Java).
 
-Discuss:
-
-\begin{itemize}
-  \item Proof carrying code
-  \item Proof assistants
-  \item Verified languages and compilers
-  \item Information flow languages
-\end{itemize}
+Another related avenue of research has been the work on proof carrying code,
+which involves a compiler producing a proof that the program satisfies some
+property along with the compiled program, allowing a smaller trusted computing
+base for security sensitive computing
+scenarios~\cite{necula2002proof,necula1998design}.  Our goal was to make a step
+towards producing a verifying compiler for a language with information flow
+types like Jif or FlowCaml, which use a type system very similar to the one
+presented in this paper for ensuring noninterference of
+programs~\cite{InfoFlowML,myers1999jflow}.
 
@@
 
-Talk about stuff we still have to do for this to be a complete project.  Talk
-about expected problems and required techniques.  Talk about what we could do
-once this project is completed and we could use it for other projects.
+\todo{Talk about stuff we still have to do for this to be a complete project.
+Talk about expected problems and required techniques.  Talk about what we could
+do once this project is completed and we could use it for other projects.}
 
@@
 
-Summarize again what we covered, where we ended up, and where we'd like to go
-next.
+\todo{Summarize again what we covered, where we ended up, and where we'd like to
+go next.}
 

### 2013-12-18 paper/paper.tex
Two columns is more comfortable to my eye for some reason...
More futzing with formatting.
Made todo's easier to pick out of paper.
Fixed some lemma labels.

@@
-\documentclass[a4paper]{article}
+\documentclass[a4paper,twocolumn]{article}
 \usepackage{mathpartir}
@@
 \usepackage{amsthm}
+\usepackage{color}
 %\usepackage{float}
@@
 \newcommand{\lift}[1]{\lfloor #1 \rfloor}
-\newcommand{\todo}[1]{\textbf{TODO: #1}}
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
   \begin{align}
@@
 
-\begin{figure}[h!]
+\begin{figure*}[t]
   Basic reductions
@@
   \label{fig:operational_semantics}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
   \begin{align*}
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
   \begin{mathpar}
@@
 
-\begin{figure}[h!]
+\begin{figure*}[t]
   \begin{mathpar}
@@
   \label{fig:value_types}
-\end{figure}
+\end{figure*}
 
@@
   \label{fig:expr_types}
-\end{figure}
+\end{figure*}
 
@@
   be stuck and this would mean that one of the projections is also stuck by
-  Lemma~\ref{lem:bracket_stuck}, a contradicton.  Thus, we know that if both
+  Lemma~\ref{lem:stuck_bracket}, a contradicton.  Thus, we know that if both
   projections step to an answer, so must the original expression.
@@
 \begin{lemma}[Substitution]
-  \label{lem:stuck_bracket}
+  \label{lem:substitution}
   Assume $\typeRule{ }{v}{s}$.  Then, $\typeRule{\Gamma[x \mapsto s]}{v'}{t}$
@@
 $\lift{e'}_i$ are identical.  Luckily, we are fairly confident, given that we
-were still able to prove Lemma~\ref{lem:bracket_stuck} despite this change, that
+were still able to prove Lemma~\ref{lem:stuck_bracket} despite this change, that
 this is not a critical error in the original paper.

### 2013-12-18 paper/paper.tex
Ugh I'm tired, I keep typing -> instead of \to.
Outlined cases for proof of subject reduction.  Wrote an initial draft of the conclusion.

@@
   \begin{align*}
-    t ::= int^l \: | \: (t -> t)^l
+    t ::= int^l \: | \: (t \to t)^l
   \end{align*}
@@
 \begin{proof}
+  By induction on the derivation of $e \to e'$.
+  \begin{case}[$\beta$]
     \todo{}
+  \end{case}
+
+  \begin{case}[if]
+    \todo{}
+  \end{case}
+
+  \begin{case}[let]
+    \todo{}
+  \end{case}
+
+  \begin{case}[lift-app]
+    \todo{}
+  \end{case}
+
+  \begin{case}[lift-if]
+    \todo{}
+  \end{case}
+
+  \begin{case}[bracket]
+    \todo{}
+  \end{case}
 \end{proof}
@@
 
-Finally, the noninteference theorem shows that if we have an expression that is
-typed with a low label, then replacing any bracketed value in the expression
-with either projection will result in the same value for the overall expression.
-In other words, if we vary any high security item in a low security expression,
-the value must not be affected.
+The noninteference theorem shows that if we have an expression that is typed
+with a low label, then replacing any bracketed value in the expression with
+either projection will result in the same value for the overall expression.  In
+other words, if we vary any high security item in a low security expression, the
+value must not be affected.
 
@@
   current assumptions reduces to $v_i$.  By Lemma~\ref{lem:completeness} we know
-  that there is an answer $a$ such that $e[x \Leftarrow v] ->^* a$.  By
+  that there is an answer $a$ such that $e[x \Leftarrow v] \to^* a$.  By
   Lemma~\ref{lem:soundness}, $\lift{a}_i = v_i'$ for all $i$ which means that
@@
 
-\todo{Summarize again what we covered, where we ended up, and where we'd like to
-go next.}
+In this paper, we present a simple information flow typed language \langName,
+prove noninterference for the language, and discuss our experiences beginning to
+mechanize this proof in coq.  Although we did not complete our proof, we were
+able to catch inaccuracies and subtleties that the proof we based our approach
+on did not handle or explicitly cover.  This is the first step towards a
+verified compiler for a language with information flow types such as Jif or
+FlowCaml, which allows for a more trustworthy compiler and language definition,
+by allowing users to verify the proofs in a proof assistant like Coq.  As far as
+we know, a verified proof of noninterference for such a language has not been
+implemented, let alone a verified compiler.  We are confident that, given more
+time and a large amount of work, such a verified compiler could be realized for
+a mature information flow language.
 

### 2013-12-18 paper/paper.tex
capital C Coq.  A specific Coq, not just any old coq.

@@
 prove noninterference for the language, and discuss our experiences beginning to
-mechanize this proof in coq.  Although we did not complete our proof, we were
+mechanize this proof in Coq.  Although we did not complete our proof, we were
 able to catch inaccuracies and subtleties that the proof we based our approach

### 2013-12-19 paper/paper.tex
Initial draft of future work section.

@@
 this could be completed however.  So far, we have proven the first few lemmas
-described in Section~\ref{sec:proof}.  This was much more difficult than the
-short natural language proofs would suggest, due to a combination of the
-stricter format required by Coq and the number of items that are elided in the
-proofs because they were considered trivial or obvious.
+described in Section~\ref{sec:proof} and defined relations for the operational
+semantics and type system as presented in Section~\ref{sec:lang}.  This was much
+more difficult than the short natural language proofs would suggest, due to a
+combination of the stricter format required by Coq and the number of items that
+are elided in the proofs because they were considered trivial or obvious.
 
@@
 
-\todo{Talk about stuff we still have to do for this to be a complete project.
-Talk about expected problems and required techniques.  Talk about what we could
-do once this project is completed and we could use it for other projects.}
+It is our hope that we could continue this work to complete a Coq proof of
+noninterference for our language.  This would require proving the lemmas and
+theorem in Section~\ref{sec:proof} along with various support lemmas and
+definitions for things that are ``obvious'' in the natural language proofs.
+It's unclear how long the remainder of the proof would take, but we are certain
+that it is just a matter of time and effort.
+
+Once a full Coq proof of noninterference exists, a next step would be to
+integrate this proof and all of the related definitions into a verified type
+checking program, which could be intergrated into a verified compiler.  We
+already know that a full verified compiler has been done for C, and this gives
+us hope that it would be possible to implement something similar for a language
+like Jif.
+
+With a verified compiler for such a language, it would be possible to implement
+a compiler that produced proofs of noninterference to be carried around with the
+compiled program.  This would be interesting in distributed settings where we
+would like to ship code around to others to run on their own machines.  There is
+already a paper on the subject of mobile code for the Fabric
+language~\cite{liu2009fabric}, a distributed computing variant of Jif,
+discussing the problem of ``mobile code'', which operates by having the client
+compile the code themselves~\cite{arden2012sharing}.  A proof-carrying code
+system could possibly remove the need for the target machine to compile the code
+themselves in order to ensure trust requirements are met.
+
+Another use of a verified compiler would be the ability to verify extensions to
+the language.  For example, the Jif and Fabric programming languages are
+actively worked on and has seen multiple extensions over the years.  Although
+it's pretty certain extensions to these languages that are published are
+correct, requiring these extensions to be added to a verified compiler would
+help ensure that the language updates will not invalidate pre-existing security
+properties of the language, because adding the extension would require adding
+code to cover the additional cases in the proofs.
+
+\todo{more.}
 

### 2013-12-19 paper/paper.tex
Most of the last proof done.

@@
   \begin{case}[$\beta$]
-    \todo{}
+    In this case, $e = (\textsf{fix} f . \lambda x . e_0) v$.  In the premises
+    of \textsc{E-App}, we have $\typeRule{ }{\textsf{fix} f . \lambda x .
+    e_0}{(t' \to t)^l}$ and $\typeRule{ }{v}{t'}$.  The derivation of the
+    function's type must end with \textsc{V-Abs} followed by some number of
+    applications of \textsc{V-Sub}.  Applying Lemma~\ref{lem:weakening} and
+    \textsc{E-Sub} to \text{V-Abs}'s premise gives $\typeRule{pc, (x \mapsto
+    t''; f \mapsto \theta)}{e'}{t}$ for some $t''$ and $\theta$ which are
+    subtypes of the function's type and the argument's type.  By \textsc{V-Sub},
+    $\typeRule{ }{v}{t''}$ and $\typeRule{ }{\textsc{fix} f . \lambda x .
+    e_0}{\theta}$.  Applying Lemma~\ref{lem:substitution} then gives
+    $\typeRule{pc}{e'}{t}$.
   \end{case}
@@
   \begin{case}[let]
-    \todo{}
+    By \textsc{E-Let} and Lemma~\ref{lem:substitution}.
   \end{case}
@@
   \begin{case}[lift-app]
-    \todo{}
+    $e = <v_1 | v_2> v$ in this case.  Then \textsc{E-App}'s premises are
+    $\typeRule{ }{<v_1 | v_2>}{(t' \to t)^l}$, $\typeRule{ }{v}{t'}$, and $l
+    \guards t$.  Lemma~\ref{lem:projection} gives $\typeRule{ }{v_i}{(t' \to
+    t)^l}$ and $\typeRule{ }{\lift{v}_i}{t'}$ for $i \in \{1, 2\}$.  Applying
+    \textsc{E-App} gives $\typeRule{pc \sqcup l}{v_i\lift{v}_i}{t}$.  Applying
+    Lemma~\ref{lem:guard} gives $l \in H$.  Since we have $l \guards t$, we can
+    apply \textsc{E-Bracket} to get $\typeRule{ }{e'}{t}$.
   \end{case}
@@
   \begin{case}[bracket]
-    \todo{}
+    In this case $e = <e_1 | e_2>$, and $e'$ is the result of taking a step in
+    one side of the brackets.  $e$'s typing derivation must end with
+    \textsc{E-Bracket}, which has premises $\typeRule{pc \sqcup pc'}{e_i}{t}$,
+    $\typeRule{pc \sqcup pc'}{e_j}{t}$, $pc' \guards t$, and $pc' \in H$.  Since
+    $pc' \in H$, we can apply the inductive hypothesis to get $\typeRule{pc
+    \sqcup pc'}{e_i'}{t}$ and $\typeRule{pc \sqcup pc'}{e_j'}{t}$.  The result
+    is then given by applying \textsc{E-Bracket}.
   \end{case}

### 2013-12-19 paper/paper.tex
Forgot an addition to the bibtex.  Finished last proof draft.

@@
   \begin{case}[if]
-    \todo{}
+    $e = \textsf{if } v \: e_1 \: e_2$ and $e' = e_i$ where $i$ is determined by
+    the value of $v$.  \textsc{E-If}'s premises include $\typeRule{pc
+    \sqcup l}{e_i}{t}$ for $i \in \{1, 2\}$ and $l \guards t$, which follows
+    from Lemma~\ref{lem:weakening}.
   \end{case}
@@
   \begin{case}[lift-if]
-    \todo{}
+    $e = \text{if } <v_1 | v_2> \: e_1 \: e_2$ and $e' = e_i$ for some $i \in
+    \{1, 2\}$ in this case.  Applying Lemma~\ref{lem:guard} to the first premise
+    of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
+    Lemma~\ref{lem:projection} to the intial premises of \textsc{E-If} and using
+    the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
+    We can then apply \textsc{E-Bracket} to get the goal.
   \end{case}

### 2013-12-19 paper/paper.tex
tweak

@@
 A reader familiar with the type system for \origLang will notice that our type
-system does not include types above the arrow in the function type, for giving
-security bounds on side effects of the function, and the type does not include
-effect labels for exceptions.  These are dropped as a result of the
+system does not include labels above the arrow in the function type, which would
+normally give security bounds on side effects of the function, nor does it
+include effect labels for exceptions.  These are dropped as a result of the
 simplifications we made in defining \langName, since there are no side effects

### 2013-12-19 paper/paper.tex

@@
 In this paper we discuss our experience trying to mechanize a proof of
-noninterference for a simple language with security labels, \langName.  This
+noninterference for a simple language with security labels, \langName, based on Pottier and Sionet's \origLang.  This
 marks the first step towards writing a verified compiler for a language with
 information flow types.  Although we did not complete the full noninterference
-proof because we ran out of time, we were able to find small inaccuracies and
-subtleties in the proofs from the paper we based our proof on.  We discuss what
-would be involved in completing the proof and how this relates to other work.
+proof, we were able to find subtle inaccuracies in the \origLang proofs.  
+We discuss what would be involved in completing the proof and how this 
+relates to other work.
 
@@
 Verified systems have been growing in popularity recently with the rise of proof
-assistants like Coq and NuPRL~\cite{coqart,nuprl}.  These systems allow a
-programmer to write programs and have proofs, which are checked and compiled by
-the system, which can assert various properties about the programs.  As we will
-discuss in Section~\ref{sec:related}, these systems have been used to formalize,
-prove correct, and implement systems that have verified properties.  This allows
-for programmers to be more confident in their languages and programs, ensuring
-that the properties they want are actually properties of their work.
+assistants like Coq and NuPRL~\cite{coqart,nuprl}.  
+These systems allow programmers to construct proofs of properties of their programs which are checked and compiled by the system.  As we will
+discuss in Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to formalize,
+prove correct, and implement systems with verified properties.
+This increases Programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).
 
-Interestingly, as far as we know there has not been work to formalize and verify
-a language, like Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow},
-that has a type system for enforcing information flow security policies.  For
-languages like these and other systems related to security, it is desirable for
-the program to be accompanied with a proof of the security properties it
-enforces.  Our goal is to make progress towards verifying the correctness of a
+The last decade has also seen a surge of research in languages which inforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
+These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.
+
+Given the growing popularity of Proof Assistants information-flow languages, we were surprised to discover 
+that there has not been work done to formalize ad verify a language with a type systems which enforces information flow security policies.
+These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties it enforces.
+Our goal is to make progress towards verifying the correctness of a
 compiler for a language with information flow types that programmers will want
@@
 
-A verified compiler for a language like Jif would require a proof of
-noninterference, which asserts that any program in the language will never
+The core theorem which distinguishes an information-flow language from a 
+``normal'' programming language is the poof of noninterference, 
+which asserts that across any two runs, every program in the language will never
 produce different outputs for low security items if only the high security input
-values are varied.  Thus, in this paper we discuss our experience working
-towards proving this theorem for a simple information flow type system as a
-first step towards creating a verified information flow compiler.
+values are varied.  It is just this theorem we wish to prove in this paper; we
+hope to learn through this effort of the unique challenges associated with
+mechanizing proofs of information-flow languages.
 
@@
 presented in Simonet and Pottier's work on information flow for an ML-like
-language, \origLang~\cite{InfoFlowML}.  This proof proceeds by designing a
-language that includes values which represent two different alternatives,
-representing the different values of high security items in two different
-executions of the program.  The proof then becomes a matter of proving that if
-the program is typed with a low security label, this does not produce a value
-that could be different between two different executions of the program.
+language, \origLang~\cite{InfoFlowML}.  The proof proceeds by designing a
+core language, extending it with a term that represents a ``pair'' of 
+values varied across two executions, and introducing these ``pairs'' of 
+different values in context where the value would have a high-security label.
+The proof then becomes a matter of showing that if
+the program is typed with a low security label, it does not terminate
+with a result that is an instance of one of these ``pairs.''
 
 In this paper we will present our simplified language, \langName,
-along with the semantics and type system (Section~\ref{sec:lang}).  We will then
-present a natural language, human readable, proof of noninterference for the
+along with its semantics and type system (Section~\ref{sec:lang}).  We will then
+present a natural language proof of noninterference for the
 language and discuss our experiences with mechanizing this proof in the Coq
@@
 of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
-limited amount of time to work on this project, we limited ourselves to a subset
+limited amount of time to work on this project, we restricted ourselves to a subset
 of the language that drops the following features:
@@
   \item References and associated memory operations.
-  \item Paired values.
+  \item Value Tuples.
   \item Exceptions.
@@
 
-While the language becomes much much less interesting for actual programming, we
+While the language becomes much more burdensome for the programmer, we
 feel that these features do not fundamentally change the problem of proving
-noninterference.  In order to mechanize a proof of noninterference for a
-language closer to \origLang, it would require simply adding mechanized proofs
+noninterference.  Mechanizing a proof of noninterference for a
+language closer to \origLang would require simply adding mechanized proofs
 for the additional cases where appropriate.

### 2013-12-19 paper/paper.tex
earlyoften

@@
 along with its semantics and type system (Section~\ref{sec:lang}).  We will then
-present a natural language proof of noninterference for the
+present a natural language, human readable, proof of noninterference for the
 language and discuss our experiences with mechanizing this proof in the Coq
@@
 of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
-limited amount of time to work on this project, we restricted ourselves to a subset
+limited amount of time to work on this project, we limited ourselves to a subset
 of the language that drops the following features:
@@
   \item References and associated memory operations.
-  \item Value Tuples.
+  \item Paired values.
   \item Exceptions.
@@
 
-While the language becomes much more burdensome for the programmer, we
+While the language becomes much much less interesting for actual programming, we
 feel that these features do not fundamentally change the problem of proving
-noninterference.  Mechanizing a proof of noninterference for a
-language closer to \origLang would require simply adding mechanized proofs
+noninterference.  In order to mechanize a proof of noninterference for a
+language closer to \origLang, it would require simply adding mechanized proofs
 for the additional cases where appropriate.
@@
 A reader familiar with the type system for \origLang will notice that our type
-system does not include labels above the arrow in the function type, which would
-normally give security bounds on side effects of the function, nor does it
-include effect labels for exceptions.  These are dropped as a result of the
+system does not include types above the arrow in the function type, for giving
+security bounds on side effects of the function, and the type does not include
+effect labels for exceptions.  These are dropped as a result of the
 simplifications we made in defining \langName, since there are no side effects

### 2013-12-19 paper/paper.tex
earlyoften

@@
 
-Furthermore, primarily due to a matter of taste, we decided to replace the
-\textsf{case} construct from \origLang for an \textsf{if} construct, which
+We also decided to replace the
+\textsf{case} construct from \origLang with an \textsf{if} construct, which
 evaluates the first expression if the test evaluates to 1 and evaluates to the
@@
 
-\langName, like \origLang, includes angle bracketed expressions, which represent
+\langName, like \origLang, includes angle bracketed expressions which represent
 values or expressions which could differ between two runs of the program.  As we

### 2013-12-19 paper/paper.tex
earlyoften

@@
 Through these changes, our language has the syntax shown in
-Figure~\ref{fig:syntax}.  As shown, final states, answers, in the language are
+Figure~\ref{fig:syntax}.  As shown, final states (answers) in the language are
 now simple values because there are no exceptions.
@@
 
-\langName has an operational semantics that should be familiar to the reader who
-has seen a single step semantics for a basic lambda calculus.  The interesting
-additions are the same as those in \origLang, where there are additional rules
+\langName has an operational semantics reminiscent of single step semantics for a basic lambda calculus.  The interesting
+additions are the same as those in \origLang -- additional rules
 for handling the bracketed values.  The reduction rules for the semantics are
@@
 states that the label $l$ is at least as restrictive as the label on the type
-$t$.  In other words, this requires that the type has at least $l$ for it's
+$t$.  In other words, this requires that the type has at least $l$ for its
 security label.  The two rules regarding this relationships for both types in

### 2013-12-19 paper/paper.tex
Started restructuring the "what we learned" section to have a list of inaccuracies found.

--- 2013-12-19 paper/paper.tex
@@
 
-We also decided to replace the
-\textsf{case} construct from \origLang with an \textsf{if} construct, which
+Furthermore, primarily due to a matter of taste, we decided to replace the
+\textsf{case} construct from \origLang for an \textsf{if} construct, which
 evaluates the first expression if the test evaluates to 1 and evaluates to the
@@
 
-\langName, like \origLang, includes angle bracketed expressions which represent
+\langName, like \origLang, includes angle bracketed expressions, which represent
 values or expressions which could differ between two runs of the program.  As we
@@
 
-Our proof of Lemma~\ref{lem:soundness} actually turned out to be instructive
-because it forced us to recognize that the original formulation in Simonet and
-Pottier's paper on \origLang~\cite{InfoFlowML} wasn't quite correct.
-Specifically, it is entirely possible that a step in the operational semantics
-does not translate to a step in the operational semantics under either
-projection, as in the case of the \ref{eq:lift-app} rule, where $\lift{e}_i$ and
-$\lift{e'}_i$ are identical.  Luckily, we are fairly confident, given that we
-were still able to prove Lemma~\ref{lem:stuck_bracket} despite this change, that
-this is not a critical error in the original paper.
+\subsection{Inaccuracies in the Original Proof}
 
-\todo{More.}
+We discovered a few small inaccuracies in \cite{InfoFlowML} by running into
+issues when formalizing the proofs in Coq.  Listed below are the primary
+corrections we were able to make.
+
+\begin{description}
+  \item[Non-nesting Brackets in Syntax]
+    \hspace{0pt}
+
+    As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
+    structure for \langName from \origLang.  Specifically, we constructed
+    values, answers, and expressions explicitly to not allow nesting of
+    bracketed values.  This was not how the syntax figures in Simonet and
+    Pottier's paper defined this, and as a result, we had trouble when our
+    proofs depended on this explicit non-nesting property that was mentioned in
+    the prose of their paper.  Realizing that our definitions did not disallow
+    this by construction after running into issues in the proof, we were able to
+    fix the definition producing what is in Figure~\ref{fig:syntax}.
+
+  \item[Correction of the Soundness Lemma]
+    \hspace{0pt}
+
+    Proving our own version of Lemma~\ref{lem:soundness} actually turned out to
+    be instructive because it forced us to recognize that the original
+    formulation in Simonet and Pottier's paper on \origLang~\cite{InfoFlowML}
+    wasn't exactly correct as stated.  Specifically, it is entirely possible
+    that a step in the operational semantics does not translate to a step in the
+    operational semantics under either projection, as in the case of the
+    \ref{eq:lift-app} rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.
+    Luckily, we are fairly confident, given that we were still able to prove
+    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
+    critical error in the original paper.
+\end{description}
 

### 2013-12-19 paper/paper.tex
earlyoften

@@
 
-    \infer[G-Int]
+    \infer[G-Abs]
     {

### 2013-12-19 paper/paper.tex
Added a todo to remind us to try to add more to the mechanizing section.

@@
     be instructive because it forced us to recognize that the original
-    formulation in Simonet and Pottier's paper on \origLang~\cite{InfoFlowML}
-    wasn't exactly correct as stated.  Specifically, it is entirely possible
-    that a step in the operational semantics does not translate to a step in the
-    operational semantics under either projection, as in the case of the
-    \ref{eq:lift-app} rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.
-    Luckily, we are fairly confident, given that we were still able to prove
+    formulation in Simonet and Pottier's paper on \origLang wasn't exactly
+    correct as stated.  Specifically, it is entirely possible that a step in the
+    operational semantics does not translate to a step in the operational
+    semantics under either projection, as in the case of the \ref{eq:lift-app}
+    rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.  Luckily, we are
+    fairly confident, given that we were still able to prove
     Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
@@
 \end{description}
+
+\todo{more?}
 

### 2013-12-19 paper/paper.tex
Sionet => Simonet

@@
 In this paper we discuss our experience trying to mechanize a proof of
-noninterference for a simple language with security labels, \langName, based on Pottier and Sionet's \origLang.  This
+noninterference for a simple language with security labels, \langName, based on
+Pottier and Simonet's \origLang.  This
 marks the first step towards writing a verified compiler for a language with

### 2013-12-19 paper/paper.tex
Typos.

@@
 
-The last decade has also seen a surge of research in languages which inforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
+The last decade has also seen a surge of research in languages which enforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
 These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.
@@
 The language's type system includes subtyping, which allows data that have a
-lower label to flow into higher labelled data.  We use the notatoin $x \leq y$
+lower label to flow into higher labelled data.  We use the notation $x \leq y$
 to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
@@
 \textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
-from the set $H$ which representes an upward closed sublattice of the labels in
+from the set $H$ which represents an upward closed sublattice of the labels in
 the system representing what is considered to be ``high'' security for the
@@
   be stuck and this would mean that one of the projections is also stuck by
-  Lemma~\ref{lem:stuck_bracket}, a contradicton.  Thus, we know that if both
+  Lemma~\ref{lem:stuck_bracket}, a contradiction.  Thus, we know that if both
   projections step to an answer, so must the original expression.
@@
     of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
-    Lemma~\ref{lem:projection} to the intial premises of \textsc{E-If} and using
+    Lemma~\ref{lem:projection} to the initial premises of \textsc{E-If} and using
     the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
@@
 
-The noninteference theorem shows that if we have an expression that is typed
+The noninterference theorem shows that if we have an expression that is typed
 with a low label, then replacing any bracketed value in the expression with
@@
 paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
-natural languae but turned out to be 40 lines of Coq code, not including all of
+natural language but turned out to be 40 lines of Coq code, not including all of
 the supporting definitions and rules.
@@
 integrate this proof and all of the related definitions into a verified type
-checking program, which could be intergrated into a verified compiler.  We
+checking program, which could be integrated into a verified compiler.  We
 already know that a full verified compiler has been done for C, and this gives

### 2013-12-19 paper/paper.tex

@@
 Figure~\ref{fig:value_types} includes typing judgements for values in the
-language, which are mostly straight forward with the exception of
+language, which are mostly straightforward with the exception of
 \textsc{V-Bracket}, which we discuss below along with the type rules for
@@
 Before discussing the task of mechanizing a noninterference proof for \langName,
-it is useful to first view what the natural language proof would look like
-first.  We first start by proving some supporting lemmas about the language's
+it is useful to first view the natural language proof.  We first start by proving some supporting lemmas about the language's
 semantics and type system, and finish with a proof of noninterference for the
@@
     \textsf{ let } x := v \textsf{ in } e$]
-    These are similar to the previous case.  The value $v$ can not be an integer
-    and so the lifted version of $v$ can not be an integer, meaning the lifted
-    expression is also stuck.
+    If $e$ is stuck and is an If-statement, then none of \ref{eq:lift-if}, \ref{eq:if-true}, and \ref{eq:if-false} applies. This means that $v$ can neither
+    be an integer or some bracketed value.  As a result, $\lift{v}_i$ cannot
+    be an integer, and so is stuck.
   \end{case}

### 2013-12-19 paper/paper.tex
Make line wrapping nicer in a long list of citations.

@@
 providing the proofs of these properties on the
-definitions~\cite{strecker2002formal,leinenbach2005towards,dave2003compiler,leroy2009formally,leroy2006formal,leroy2009formal,leroy2012compcert,chlipala2010verified,berghofer2004extracting}.
-This has been a popular area, which is primarily focused on more mainstream
-languages and compilers (such as C and Java).
+definitions~\cite{strecker2002formal, leinenbach2005towards, dave2003compiler,
+leroy2009formally, leroy2006formal, leroy2009formal, leroy2012compcert,
+chlipala2010verified, berghofer2004extracting}.  This has been a popular area,
+which is primarily focused on more mainstream languages and compilers (such as C
+and Java).
 

### 2013-12-19 paper/paper.tex
expand if proof in lemma2

--- 2013-12-19 paper/paper.tex
@@
 In this paper we discuss our experience trying to mechanize a proof of
-noninterference for a simple language with security labels, \langName, based on
-Pottier and Simonet's \origLang.  This
+noninterference for a simple language with security labels, \langName, based on Pottier and Sionet's \origLang.  This
 marks the first step towards writing a verified compiler for a language with
@@
 
-The last decade has also seen a surge of research in languages which enforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
+The last decade has also seen a surge of research in languages which inforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
 These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.
@@
 The language's type system includes subtyping, which allows data that have a
-lower label to flow into higher labelled data.  We use the notation $x \leq y$
+lower label to flow into higher labelled data.  We use the notatoin $x \leq y$
 to denote that $x$ is a subtype of $y$, whether both $x$ and $y$ are labels or
@@
 \textsc{V-Bracket} and \textsc{E-Bracket} both require that there exists a label
-from the set $H$ which represents an upward closed sublattice of the labels in
+from the set $H$ which representes an upward closed sublattice of the labels in
 the system representing what is considered to be ``high'' security for the
@@
   be stuck and this would mean that one of the projections is also stuck by
-  Lemma~\ref{lem:stuck_bracket}, a contradiction.  Thus, we know that if both
+  Lemma~\ref{lem:stuck_bracket}, a contradicton.  Thus, we know that if both
   projections step to an answer, so must the original expression.
@@
     of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
-    Lemma~\ref{lem:projection} to the initial premises of \textsc{E-If} and using
+    Lemma~\ref{lem:projection} to the intial premises of \textsc{E-If} and using
     the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
@@
 
-The noninterference theorem shows that if we have an expression that is typed
+The noninteference theorem shows that if we have an expression that is typed
 with a low label, then replacing any bracketed value in the expression with
@@
 paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
-natural language but turned out to be 40 lines of Coq code, not including all of
+natural languae but turned out to be 40 lines of Coq code, not including all of
 the supporting definitions and rules.
 
-\subsection{Inaccuracies in the Original Proof}
+Our proof of Lemma~\ref{lem:soundness} actually turned out to be instructive
+because it forced us to recognize that the original formulation in Simonet and
+Pottier's paper on \origLang~\cite{InfoFlowML} wasn't quite correct.
+Specifically, it is entirely possible that a step in the operational semantics
+does not translate to a step in the operational semantics under either
+projection, as in the case of the \ref{eq:lift-app} rule, where $\lift{e}_i$ and
+$\lift{e'}_i$ are identical.  Luckily, we are fairly confident, given that we
+were still able to prove Lemma~\ref{lem:stuck_bracket} despite this change, that
+this is not a critical error in the original paper.
 
-We discovered a few small inaccuracies in \cite{InfoFlowML} by running into
-issues when formalizing the proofs in Coq.  Listed below are the primary
-corrections we were able to make.
-
-\begin{description}
-  \item[Non-nesting Brackets in Syntax]
-    \hspace{0pt}
-
-    As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
-    structure for \langName from \origLang.  Specifically, we constructed
-    values, answers, and expressions explicitly to not allow nesting of
-    bracketed values.  This was not how the syntax figures in Simonet and
-    Pottier's paper defined this, and as a result, we had trouble when our
-    proofs depended on this explicit non-nesting property that was mentioned in
-    the prose of their paper.  Realizing that our definitions did not disallow
-    this by construction after running into issues in the proof, we were able to
-    fix the definition producing what is in Figure~\ref{fig:syntax}.
-
-  \item[Correction of the Soundness Lemma]
-    \hspace{0pt}
-
-    Proving our own version of Lemma~\ref{lem:soundness} actually turned out to
-    be instructive because it forced us to recognize that the original
-    formulation in Simonet and Pottier's paper on \origLang wasn't exactly
-    correct as stated.  Specifically, it is entirely possible that a step in the
-    operational semantics does not translate to a step in the operational
-    semantics under either projection, as in the case of the \ref{eq:lift-app}
-    rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.  Luckily, we are
-    fairly confident, given that we were still able to prove
-    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
-    critical error in the original paper.
-\end{description}
-
-\todo{more?}
+\todo{More.}
 
@@
 integrate this proof and all of the related definitions into a verified type
-checking program, which could be integrated into a verified compiler.  We
+checking program, which could be intergrated into a verified compiler.  We
 already know that a full verified compiler has been done for C, and this gives

### 2013-12-19 paper/paper.tex

@@
 
-Our proof of Lemma~\ref{lem:soundness} actually turned out to be instructive
-because it forced us to recognize that the original formulation in Simonet and
-Pottier's paper on \origLang~\cite{InfoFlowML} wasn't quite correct.
-Specifically, it is entirely possible that a step in the operational semantics
-does not translate to a step in the operational semantics under either
-projection, as in the case of the \ref{eq:lift-app} rule, where $\lift{e}_i$ and
-$\lift{e'}_i$ are identical.  Luckily, we are fairly confident, given that we
-were still able to prove Lemma~\ref{lem:stuck_bracket} despite this change, that
-this is not a critical error in the original paper.
+\subsection{Inaccuracies in the Original Proof}
 
-\todo{More.}
+We discovered a few small inaccuracies in \cite{InfoFlowML} by running into
+issues when formalizing the proofs in Coq.  Listed below are the primary
+corrections we were able to make.
+
+\begin{description}
+  \item[Non-nesting Brackets in Syntax]
+    \hspace{0pt}
+
+    As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
+    structure for \langName from \origLang.  Specifically, we constructed
+    values, answers, and expressions explicitly to not allow nesting of
+    bracketed values.  This was not how the syntax figures in Simonet and
+    Pottier's paper defined this, and as a result, we had trouble when our
+    proofs depended on this explicit non-nesting property that was mentioned in
+    the prose of their paper.  Realizing that our definitions did not disallow
+    this by construction after running into issues in the proof, we were able to
+    fix the definition producing what is in Figure~\ref{fig:syntax}.
+
+  \item[Correction of the Soundness Lemma]
+    \hspace{0pt}
+
+    Proving our own version of Lemma~\ref{lem:soundness} actually turned out to
+    be instructive because it forced us to recognize that the original
+    formulation in Simonet and Pottier's paper on \origLang wasn't exactly
+    correct as stated.  Specifically, it is entirely possible that a step in the
+    operational semantics does not translate to a step in the operational
+    semantics under either projection, as in the case of the \ref{eq:lift-app}
+    rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.  Luckily, we are
+    fairly confident, given that we were still able to prove
+    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
+    critical error in the original paper.
+\end{description}
+
+\todo{more?}
 

### 2013-12-19 paper/paper.tex
Truthifying the abstract.
subclauses usually need a subject.

--- 2013-12-19 paper/paper.tex
@@
 information flow types.  Although we did not complete the full noninterference
-proof, we were able to find subtle inaccuracies in the \origLang proofs.  
-We discuss what would be involved in completing the proof and how this 
-relates to other work.
+proof, we were able to find subtle inaccuracies in \origLang's original
+formulation and theorems.  We discuss what would be involved in completing the
+proof and how this relates to other work.
 
@@
     be an integer or some bracketed value.  As a result, $\lift{v}_i$ cannot
-    be an integer, and so is stuck.
+    be an integer, and so it is stuck.
   \end{case}

### 2013-12-19 paper/paper.tex
some additions to the mechanizing proofs section

--- 2013-12-19 paper/paper.tex
@@
 As mentioned above, we unfortunately were unable to complete our noninterference
-proof in Coq by the time of this report.  Given more time, we are confident that
-this could be completed however.  So far, we have proven the first few lemmas
-described in Section~\ref{sec:proof} and defined relations for the operational
-semantics and type system as presented in Section~\ref{sec:lang}.  This was much
-more difficult than the short natural language proofs would suggest, due to a
-combination of the stricter format required by Coq and the number of items that
-are elided in the proofs because they were considered trivial or obvious.
+proof in Coq by the time of this report.
+As of this writing, we have written definitions of a small-step operational semantics, a big-step operational semantics, typing judgments, and typing derivations for \langName.
+We have proven a few lemmas from section~\ref{sec:proof} relating to the operations semantics (including soundness), but remain mired in complexity on the remainder of the mechanization effort.
 
-We found that the process of mechanizing the lemmas that we proved was a useful
-exercise in checking our understanding of our language, statements, and the work
-of Simonet and Pottier which we based our approach off of.  We found ourselves
-often surprised by the number of items that were elided by the proofs in the
-paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
-natural languae but turned out to be 40 lines of Coq code, not including all of
-the supporting definitions and rules.
+In the process of mechanizing these lemmas, we empirically observed that the process of translating English into Coq can involve a nearly fifty-fold increase in the size of even the most basic proofs.
+While this was partially a natural effect of writing proofs in Coq, we found that much of the complexity stemmed from handling cases elided in the English-Language proofs; and in these cases we found hidden subtleties which appear to have escaped even Simonet and Pottier.
 
@@
 
-We discovered a few small inaccuracies in \cite{InfoFlowML} by running into
-issues when formalizing the proofs in Coq.  Listed below are the primary
+We discovered a few small inaccuracies in \cite{InfoFlowML} during the 
+process of formalizing our proofs in Coq.  Listed below are the primary
 corrections we were able to make.
+
+While some inaccuracies did invalidate Simonet and Pottier's original
+Lemmas, we universally found that a slightly different formulation was
+provable and remained usable in future proofs.
 
@@
     As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
-    structure for \langName from \origLang.  Specifically, we constructed
+    structure for \langName ~than was seen in~ \origLang.  Specifically, we constructed
     values, answers, and expressions explicitly to not allow nesting of
-    bracketed values.  This was not how the syntax figures in Simonet and
-    Pottier's paper defined this, and as a result, we had trouble when our
-    proofs depended on this explicit non-nesting property that was mentioned in
-    the prose of their paper.  Realizing that our definitions did not disallow
+    bracketed values.  The syntax defined by Simonet and Pottier did allow
+    for this nesting, and authors dismissed is as a non-sensible
+    construction.  Unfortunately, allowing such syntactic nesting 
+    prevents the completeness Lemma from being proveable.
+    Realizing that our definitions did not disallow
     this by construction after running into issues in the proof, we were able to
-    fix the definition producing what is in Figure~\ref{fig:syntax}.
+    fix the definition, producing what is in Figure~\ref{fig:syntax}.
 
@@
 
-    Proving our own version of Lemma~\ref{lem:soundness} actually turned out to
-    be instructive because it forced us to recognize that the original
-    formulation in Simonet and Pottier's paper on \origLang wasn't exactly
-    correct as stated.  Specifically, it is entirely possible that a step in the
-    operational semantics does not translate to a step in the operational
-    semantics under either projection, as in the case of the \ref{eq:lift-app}
-    rule, where $\lift{e}_i$ and $\lift{e'}_i$ are identical.  Luckily, we are
+    Proving our own version of Lemma~\ref{lem:soundness} was uniquely
+    instructive because it forced us to recognize that the original
+    formulation in Simonet and Pottier's paper on \origLang ~wasn't 
+    correct as stated.  Simonet and Pottier's Lemma asserts that 
+    if $e \rightarrow e'$, then $\lift{e}_i \rightarrow \lift{e'}_i$.
+    This does not hold in the case where $e$ is of the form $<v1 | v2> v$;
+    In this case, the single reduction rule that applies is \ref{eq:lift-app},
+    which does not step $\lift{e}_i \rightarrow \lift{e'}_i$ for any $i$ 
+    (as $\lift{e}_i$ and $\lift{e'}_i$ are identical).  Luckily, we are
     fairly confident, given that we were still able to prove
@@
 \end{description}
-
-\todo{more?}
 

### 2013-12-19 paper/paper.tex

--- 2013-12-19 paper/paper.tex
@@
 As mentioned above, we unfortunately were unable to complete our noninterference
-proof in Coq by the time of this report.  Given more time, we are confident that
-this could be completed however.  So far, we have proven the first few lemmas
-described in Section~\ref{sec:proof} and defined relations for the operational
-semantics and type system as presented in Section~\ref{sec:lang}.  This was much
-more difficult than the short natural language proofs would suggest, due to a
-combination of the stricter format required by Coq and the number of items that
-are elided in the proofs because they were considered trivial or obvious.
+proof in Coq by the time of this report.
+As of this writing, we have written definitions of a small-step operational semantics, a big-step operational semantics, typing judgments, and typing derivations for \langName.
+We have proven a few lemmas from section~\ref{sec:proof} relating to the operations semantics (including soundness), but remain mired in complexity on the remainder of the mechanization effort.
 
-We found that the process of mechanizing the lemmas that we proved was a useful
-exercise in checking our understanding of our language, statements, and the work
-of Simonet and Pottier which we based our approach off of.  We found ourselves
+In the process of mechanizing these lemmas, we empirically observed that the process of translating English into Coq can involve a nearly fifty-fold increase in the size of even the most basic proofs.
+While this was partially a natural effect of writing proofs in Coq, we found that much of the complexity stemmed from handling cases elided in the English-Language proofs; and in these cases we found hidden subtleties which appear to have escaped even Simonet and Pottier.
+We found ourselves
 often surprised by the number of items that were elided by the proofs in the

### 2013-12-19 paper/paper.tex
it's certain THAT something is true.

@@
 actively worked on and has seen multiple extensions over the years.  Although
-it's pretty certain extensions to these languages that are published are
+it's pretty certain that extensions to these languages that are published are
 correct, requiring these extensions to be added to a verified compiler would

### 2013-12-19 paper/paper.tex
CERTAIN THAT.
THAT.

THAAAAAAAAAAAAAAAAAAAAAAAAT.

--- 2013-12-19 paper/paper.tex
@@
 
+Once our noninterference proof has been sucessfully mechanized for \langName, we will incrementally enhance the features of \langName, adding in absent constructs such as bind, references, and exceptions.
+We will then re-state and re-prove the noninterference result in a language very much like \origLang.
+
 Once a full Coq proof of noninterference exists, a next step would be to
@@
 us hope that it would be possible to implement something similar for a language
-like Jif.
+like Jif.  We are aware that any such effort would require many person-years of effort to complete, but we remain convinced that such an effort would be a worthwhile initiative.
 
-With a verified compiler for such a language, it would be possible to implement
+With our verification effort complete, it would be possible to implement
 a compiler that produced proofs of noninterference to be carried around with the
@@
 the language.  For example, the Jif and Fabric programming languages are
-actively worked on and has seen multiple extensions over the years.  Although
-it's pretty certain extensions to these languages that are published are
-correct, requiring these extensions to be added to a verified compiler would
-help ensure that the language updates will not invalidate pre-existing security
-properties of the language, because adding the extension would require adding
-code to cover the additional cases in the proofs.
-
-\todo{more.}
+actively worked on and have seen multiple extensions over the years.  Although
+it's pretty certain that extensions to these languages that are published are
+correct, errors to creep into even the most scrutinized of paper results (as we
+have observed in this project), and so requiring these extensions to come with
+mechanized proofs of correctness before incorporation into a verified compiler
+would help ensure that the language updates will not invalidate pre-existing
+security properties of the language.
 

### 2013-12-19 paper/paper.tex
content additions largely complete, editing time

@@
 actively worked on and have seen multiple extensions over the years.  Although
-it's pretty certain that extensions to these languages that are published are
-correct, errors to creep into even the most scrutinized of paper results (as we
-have observed in this project), and so requiring these extensions to come with
-mechanized proofs of correctness before incorporation into a verified compiler
-would help ensure that the language updates will not invalidate pre-existing
-security properties of the language.
+it's pretty certain extensions to these languages that are published are
+correct, errors to creep into even the most scrutinized of paper results (as we have observed in this project), and so requiring these extensions to come with mechanized proofs of correctness before incorporation into a verified compiler would
+help ensure that the language updates will not invalidate pre-existing security
+properties of the language.
 

### 2013-12-19 paper/paper.tex
Abstract tweak.
more copyediting

@@
 prove correct, and implement systems with verified properties.
-This increases Programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).
+This increases programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).
 
@@
 Given the growing popularity of Proof Assistants information-flow languages, we were surprised to discover 
-that there has not been work done to formalize ad verify a language with a type systems which enforces information flow security policies.
-These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties it enforces.
+that there has not been work done to formalize ad verify a language with a type system which enforces information flow security policies.
+These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties they enforce.
 Our goal is to make progress towards verifying the correctness of a
@@
 produce different outputs for low security items if only the high security input
-values are varied.  It is just this theorem we wish to prove in this paper; we
+values are varied.  It is this theorem we wish to prove in this paper; we
 hope to learn through this effort of the unique challenges associated with
@@
 
-Our language, \langName,  and proof are a simplified version of what is
+Our language and proof, \langName, are a simplified version of what is
 presented in Simonet and Pottier's work on information flow for an ML-like
@@
 values varied across two executions, and introducing these ``pairs'' of 
-different values in context where the value would have a high-security label.
+values in contexts where the value would have a high-security label.
 The proof then becomes a matter of showing that if
@@
   \item References and associated memory operations.
-  \item Value Tuples.
+  \item Value 2-Tuples.
   \item Exceptions.
@@
 
-While the language becomes much more burdensome for the programmer, we
-feel that these features do not fundamentally change the problem of proving
+While programming in this language becomes much more burdensome, we
+feel that the removal of these features does not fundamentally 
+change the problem of proving
 noninterference.  Mechanizing a proof of noninterference for a
@@
 it's pretty certain extensions to these languages that are published are
-correct, errors to creep into even the most scrutinized of paper results (as we have observed in this project), and so requiring these extensions to come with mechanized proofs of correctness before incorporation into a verified compiler would
-help ensure that the language updates will not invalidate pre-existing security
+correct, errors to creep into even the most scrutinized of paper results (as we have observed in this project), and so requiring these extensions to come with mechanized proofs of correctness before incorporation into a verified compiler would help ensure that the language updates will not invalidate pre-existing security
 properties of the language.

### 2013-12-19 paper/paper.tex

@@
 proof, we were able to find subtle inaccuracies in \origLang's original
-formulation and theorems.  We discuss what would be involved in completing the
-proof and how this relates to other work.
+formulation and theorems.  Furthermore, we discuss what would be involved in
+completing the proof and how this relates to other work.
 
@@
 actively worked on and have seen multiple extensions over the years.  Although
-it's pretty certain extensions to these languages that are published are
-correct, errors to creep into even the most scrutinized of paper results (as we have observed in this project), and so requiring these extensions to come with mechanized proofs of correctness before incorporation into a verified compiler would help ensure that the language updates will not invalidate pre-existing security
-properties of the language.
+it's pretty certain that extensions to these languages that are published are
+correct, errors to creep into even the most scrutinized of paper results (as we
+have observed in this project), and so requiring these extensions to come with
+mechanized proofs of correctness before incorporation into a verified compiler
+would help ensure that the language updates will not invalidate pre-existing
+security properties of the language.
 

### 2013-12-19 paper/paper.tex
more copyediting

@@
 noninterference.  Mechanizing a proof of noninterference for a
-language closer to \origLang would require simply adding mechanized proofs
+language closer to \origLang would require adding mechanized proofs
 for the additional cases where appropriate.
@@
 evaluates the first expression if the test evaluates to 1 and evaluates to the
-second expression otherwise.
+second expression otherwise.  This simplifies our proof structure by eliminating
+one case of capture-avoiding substitution.
 

### 2013-12-19 paper/paper.tex
more copyediting

@@
 \langName, like \origLang, includes angle bracketed expressions which represent
-values or expressions which could differ between two runs of the program.  As we
+values and expressions which could differ between two runs of the program.  As we
 will see when discussing the type system, this corresponds to values and

### 2013-12-19 paper/paper.tex
I think I fixed syntax.

@@
     v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag*{ } \\
-    v' & ::= v \: | \: <v | v> \tag{values}\\
+    v' & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e' \: | \: <v | v> \tag{values}\\
     a & ::= v \tag*{ } \\
@@
     } x := v \textsf{ in } e \tag*{ } \\
-    e' & ::= e \: | \: <e | e> \tag{expressions}
+    e' & ::= a' \: | \: v' v' \: | \: \textsf{if } v' \: e' \: e' \: | \:
+    \textsf{let } x := v' \textsf{ in } e' \: | \: <e | e> \tag{expressions}
   \end{align}

### 2013-12-19 paper/paper.tex

@@
 
-\newcommand{\langName}[0]{TinyML\textsuperscript{2}}
-\newcommand{\origLang}[0]{CoreML\textsuperscript{2}}
+\newcommand{\langName}[0]{TinyML\textsuperscript{2}~}
+\newcommand{\origLang}[0]{CoreML\textsuperscript{2}~}
 \newcommand{\typeRule}[3]{#1 \vdash #2 \colon #3}
@@
 presented in Simonet and Pottier's work on information flow for an ML-like
-language, \origLang~\cite{InfoFlowML}.  The proof proceeds by designing a
+language, \origLang \cite{InfoFlowML}.  The proof proceeds by designing a
 core language, extending it with a term that represents a ``pair'' of 
@@
 Figure~\ref{fig:syntax}.  As shown, final states (answers) in the language are
-now simple values because there are no exceptions.
+now simple values.
 
@@
     As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
-    structure for \langName ~than was seen in~ \origLang.  Specifically, we constructed
+    structure for \langName than was seen in~ \origLang.  Specifically, we constructed
     values, answers, and expressions explicitly to not allow nesting of
@@
     instructive because it forced us to recognize that the original
-    formulation in Simonet and Pottier's paper on \origLang ~wasn't 
+    formulation in Simonet and Pottier's paper on \origLang wasn't 
     correct as stated.  Simonet and Pottier's Lemma asserts that 

### 2013-12-19 paper/paper.tex
tilde tests

@@
     v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag*{ } \\
-    v' & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e' \: | \: <v | v> \tag{values}\\
+    v' & ::= v \: | \: <v | v> \tag{values}\\
     a & ::= v \tag*{ } \\
@@
     } x := v \textsf{ in } e \tag*{ } \\
-    e' & ::= a' \: | \: v' v' \: | \: \textsf{if } v' \: e' \: e' \: | \:
-    \textsf{let } x := v' \textsf{ in } e' \: | \: <e | e> \tag{expressions}
+    e' & ::= e \: | \: <e | e> \tag{expressions}
   \end{align}

### 2013-12-19 paper/paper.tex
tilde tests

--- 2013-12-19 paper/paper.tex
@@
 
-\langName has a very simple type system, including only integers and lambda
-abstractions, which include fixpoints by default, along with appropriate
+\langName has a very simple type system. It consists of only integers and lambda
+abstractions (which include fixpoints by default) along with appropriate
 security labels for both, as shown in Figure~\ref{fig:types}.  In \langName,
-security labels specify confidentiality policies, specifying what other data the
+security labels denote confidentiality policies, which specify what other data the
 labelled data can affect.\footnote{It is not difficult to imagine how to extend
@@
 be affected by items at the same label or lower (so high data can not affect low
-labelled data).
+data).
 
@@
 A reader familiar with the type system for \origLang will notice that our type
-system does not include labels above the arrow in the function type, which would
-normally give security bounds on side effects of the function, nor does it
+system does not include labels above the arrow in the function type (which would
+normally give security bounds on side effects of the function) nor does it
 include effect labels for exceptions.  These are dropped as a result of the
@@
 $t$.  In other words, this requires that the type has at least $l$ for its
-security label.  The two rules regarding this relationships for both types in
+security label.  The two rules regarding this relationship for both types in
 our language are given in Figure~\ref{fig:guard_rules}.
@@
 Figure~\ref{fig:value_types} includes typing judgements for values in the
-language, which are mostly straightforward with the exception of
-\textsc{V-Bracket}, which we discuss below along with the type rules for
-expressions.
+language, which are mostly straightforward. The exception is
+\textsc{V-Bracket}, which we discuss below.
 

### 2013-12-19 paper/paper.tex
Line wrapping.

@@
 noninterference for a simple language with security labels, \langName, based on
-Pottier and Simonet's \origLang.  This
-marks the first step towards writing a verified compiler for a language with
-information flow types.  Although we did not complete the full noninterference
-proof, we were able to find subtle inaccuracies in \origLang's original
-formulation and theorems.  Furthermore, we discuss what would be involved in
-completing the proof and how this relates to other work.
+Pottier and Simonet's \origLang.  This marks the first step towards writing a
+verified compiler for a language with information flow types.  Although we did
+not complete the full noninterference proof, we were able to find subtle
+inaccuracies in \origLang's original formulation and theorems.  Furthermore, we
+discuss what would be involved in completing the proof and how this relates to
+other work.
 
@@
 Verified systems have been growing in popularity recently with the rise of proof
-assistants like Coq and NuPRL~\cite{coqart,nuprl}.  
-These systems allow programmers to construct proofs of properties of their programs which are checked and compiled by the system.  As we will
-discuss in Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to formalize,
-prove correct, and implement systems with verified properties.
-This increases programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).
+assistants like Coq and NuPRL~\cite{coqart,nuprl}.  These systems allow
+programmers to construct proofs of properties of their programs which are
+checked and compiled by the system.  As we will discuss in
+Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to
+formalize, prove correct, and implement systems with verified properties.  This
+increases programmers' confidence in their languages and programs; to prove
+correctness of a function is a much stronger guarantee than simply testing
+functionality (or asserting correctness in the absence of tests).
 
-The last decade has also seen a surge of research in languages which enforce information flow security, such as Jif~\cite{myers1999jflow} or FlowCaml~\cite{simonet2003flow}.
-These languages provide features which programmers can use to enforce integrity and confidentiality of data as it is used throughout computations.
+The last decade has also seen a surge of research in languages which enforce
+information flow security, such as Jif~\cite{myers1999jflow} or
+FlowCaml~\cite{simonet2003flow}.  These languages provide features which
+programmers can use to enforce integrity and confidentiality of data as it is
+used throughout computations.
 
-Given the growing popularity of Proof Assistants information-flow languages, we were surprised to discover 
-that there has not been work done to formalize ad verify a language with a type system which enforces information flow security policies.
-These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties they enforce.
-Our goal is to make progress towards verifying the correctness of a
-compiler for a language with information flow types that programmers will want
-to use.
+Given the growing popularity of Proof Assistants information-flow languages, we
+were surprised to discover that there has not been work done to formalize ad
+verify a language with a type system which enforces information flow security
+policies.  These languages seem a natural choice for formal verification, as it
+is desirable for programs written in these languages to be accompanied with a
+proof of the security properties they enforce.  Our goal is to make progress
+towards verifying the correctness of a compiler for a language with information
+flow types that programmers will want to use.
 
 The core theorem which distinguishes an information-flow language from a
-``normal'' programming language is the poof of noninterference, 
-which asserts that across any two runs, every program in the language will never
-produce different outputs for low security items if only the high security input
-values are varied.  It is this theorem we wish to prove in this paper; we
-hope to learn through this effort of the unique challenges associated with
-mechanizing proofs of information-flow languages.
+``normal'' programming language is the poof of noninterference, which asserts
+that across any two runs, every program in the language will never produce
+different outputs for low security items if only the high security input values
+are varied.  It is this theorem we wish to prove in this paper; we hope to learn
+through this effort of the unique challenges associated with mechanizing proofs
+of information-flow languages.
 
-Our language and proof, \langName, are a simplified version of what is
-presented in Simonet and Pottier's work on information flow for an ML-like
-language, \origLang \cite{InfoFlowML}.  The proof proceeds by designing a
-core language, extending it with a term that represents a ``pair'' of 
-values varied across two executions, and introducing these ``pairs'' of 
-values in contexts where the value would have a high-security label.
-The proof then becomes a matter of showing that if
-the program is typed with a low security label, it does not terminate
-with a result that is an instance of one of these ``pairs.''
+Our language and proof, \langName, are a simplified version of what is presented
+in Simonet and Pottier's work on information flow for an ML-like language,
+\origLang \cite{InfoFlowML}.  The proof proceeds by designing a core language,
+extending it with a term that represents a ``pair'' of values varied across two
+executions, and introducing these ``pairs'' of values in contexts where the
+value would have a high-security label.  The proof then becomes a matter of
+showing that if the program is typed with a low security label, it does not
+terminate with a result that is an instance of one of these ``pairs.''
 
-In this paper we will present our simplified language, \langName,
-along with its semantics and type system (Section~\ref{sec:lang}).  We will then
-present a natural language proof of noninterference for the
-language and discuss our experiences with mechanizing this proof in the Coq
-proof assistant (Sections~\ref{sec:proof} and \ref{sec:mechanizing}).  Finally,
-we will discuss related work, next steps, and conclude
-(Sections~\ref{sec:related}, \ref{sec:future}, and \ref{sec:conclusion}).
+In this paper we will present our simplified language, \langName, along with its
+semantics and type system (Section~\ref{sec:lang}).  We will then present a
+natural language proof of noninterference for the language and discuss our
+experiences with mechanizing this proof in the Coq proof assistant
+(Sections~\ref{sec:proof} and \ref{sec:mechanizing}).  Finally, we will discuss
+related work, next steps, and conclude (Sections~\ref{sec:related},
+\ref{sec:future}, and \ref{sec:conclusion}).
 
@@
 of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
-limited amount of time to work on this project, we restricted ourselves to a subset
-of the language that drops the following features:
+limited amount of time to work on this project, we restricted ourselves to a
+subset of the language that drops the following features:
 
@@
 
-While programming in this language becomes much more burdensome, we
-feel that the removal of these features does not fundamentally 
-change the problem of proving
-noninterference.  Mechanizing a proof of noninterference for a
-language closer to \origLang would require adding mechanized proofs
-for the additional cases where appropriate.
+While programming in this language becomes much more burdensome, we feel that
+the removal of these features does not fundamentally change the problem of
+proving noninterference.  Mechanizing a proof of noninterference for a language
+closer to \origLang would require adding mechanized proofs for the additional
+cases where appropriate.
 
-We also decided to replace the
-\textsf{case} construct from \origLang with an \textsf{if} construct, which
-evaluates the first expression if the test evaluates to 1 and evaluates to the
-second expression otherwise.  This simplifies our proof structure by eliminating
-one case of capture-avoiding substitution.
+We also decided to replace the \textsf{case} construct from \origLang with an
+\textsf{if} construct, which evaluates the first expression if the test
+evaluates to 1 and evaluates to the second expression otherwise.  This
+simplifies our proof structure by eliminating one case of capture-avoiding
+substitution.
 
 \langName, like \origLang, includes angle bracketed expressions which represent
-values and expressions which could differ between two runs of the program.  As we
-will see when discussing the type system, this corresponds to values and
+values and expressions which could differ between two runs of the program.  As
+we will see when discussing the type system, this corresponds to values and
 expressions that have high security labels on their types.
@@
     v & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e \tag*{ } \\
-    v' & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e' \: | \: <v | v> \tag{values}\\
+    v' & ::= x \: | \: k \: | \: \textsf{fix} f . \lambda x . e' \: | \: <v | v>
+    \tag{values}\\
     a & ::= v \tag*{ } \\
@@
 
-\langName has an operational semantics reminiscent of single step semantics for a basic lambda calculus.  The interesting
-additions are the same as those in \origLang -- additional rules
-for handling the bracketed values.  The reduction rules for the semantics are
-given in Figure~\ref{fig:operational_semantics}.
+\langName has an operational semantics reminiscent of single step semantics for
+a basic lambda calculus.  The interesting additions are the same as those in
+\origLang -- additional rules for handling the bracketed values.  The reduction
+rules for the semantics are given in Figure~\ref{fig:operational_semantics}.
 
@@
 security labels for both, as shown in Figure~\ref{fig:types}.  In \langName,
-security labels denote confidentiality policies, which specify what other data the
-labelled data can affect.\footnote{It is not difficult to imagine how to extend
-our system to include integrity labels.  However, for the sake of simplicity, we
-decided to concern ourselves with only confidentiality at first}  Data can only
-be affected by items at the same label or lower (so high data can not affect low
-data).
+security labels denote confidentiality policies, which specify what other data
+the labelled data can affect.\footnote{It is not difficult to imagine how to
+extend our system to include integrity labels.  However, for the sake of
+simplicity, we decided to concern ourselves with only confidentiality at first}
+Data can only be affected by items at the same label or lower (so high data can
+not affect low data).
 
@@
 $t$.  In other words, this requires that the type has at least $l$ for its
-security label.  The two rules regarding this relationship for both types in
-our language are given in Figure~\ref{fig:guard_rules}.
+security label.  The two rules regarding this relationship for both types in our
+language are given in Figure~\ref{fig:guard_rules}.
 
@@
 Figure~\ref{fig:value_types} includes typing judgements for values in the
-language, which are mostly straightforward. The exception is
-\textsc{V-Bracket}, which we discuss below.
+language, which are mostly straightforward. The exception is \textsc{V-Bracket},
+which we discuss below.
 
@@
 Before discussing the task of mechanizing a noninterference proof for \langName,
-it is useful to first view the natural language proof.  We first start by proving some supporting lemmas about the language's
-semantics and type system, and finish with a proof of noninterference for the
-language.  Since \langName is a stripped down version of \origLang, the proofs
-will look like a simplified version of the proofs found in Pottier and Simonet's
-paper~\cite{InfoFlowML}.
+it is useful to first view the natural language proof.  We first start by
+proving some supporting lemmas about the language's semantics and type system,
+and finish with a proof of noninterference for the language.  Since \langName is
+a stripped down version of \origLang, the proofs will look like a simplified
+version of the proofs found in Pottier and Simonet's paper~\cite{InfoFlowML}.
 
@@
   \begin{case}[$e = v_1 v_2$]
-    If $e$ is stuck and is an application, this means that neither
-    \ref{eq:beta} nor \ref{eq:lift-app} applies.  This means that $v_1$ can
-    neither be a fixpoint function or some bracketed value.  As a result,
-    $\lift{v_1}_i$ can not be a fixpoint function, and so it is stuck.
+    If $e$ is stuck and is an application, this means that neither \ref{eq:beta}
+    nor \ref{eq:lift-app} applies.  This means that $v_1$ can neither be a
+    fixpoint function or some bracketed value.  As a result, $\lift{v_1}_i$ can
+    not be a fixpoint function, and so it is stuck.
   \end{case}
 
-  \begin{case}[$e = \textsf{ if } v \: e_1 \: e_2$ or $e =
-    \textsf{ let } x := v \textsf{ in } e$]
-    If $e$ is stuck and is an If-statement, then none of \ref{eq:lift-if}, \ref{eq:if-true}, and \ref{eq:if-false} applies. This means that $v$ can neither
-    be an integer or some bracketed value.  As a result, $\lift{v}_i$ cannot
-    be an integer, and so it is stuck.
+  \begin{case}[$e = \textsf{ if } v \: e_1 \: e_2$ or $e = \textsf{ let } x := v
+    \textsf{ in } e$]
+    If $e$ is stuck and is an If-statement, then none of \ref{eq:lift-if},
+    \ref{eq:if-true}, and \ref{eq:if-false} applies. This means that $v$ can
+    neither be an integer or some bracketed value.  As a result, $\lift{v}_i$
+    cannot be an integer, and so it is stuck.
   \end{case}
@@
 
-    Otherwise, we are in the case where $v'$ is some variable $y \neq x$.
-    Then, given that $v'[x \Leftarrow v] = v' = y$ and since $\Gamma(y) =
-    \Gamma[x \mapsto s]$, we can see that $\typeRule{\Gamma[x \mapsto
-    s]}{v'}{t}$ implies that $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.
+    Otherwise, we are in the case where $v'$ is some variable $y \neq x$.  Then,
+    given that $v'[x \Leftarrow v] = v' = y$ and since $\Gamma(y) = \Gamma[x
+    \mapsto s]$, we can see that $\typeRule{\Gamma[x \mapsto s]}{v'}{t}$ implies
+    that $\typeRule{\Gamma}{v'[x \Leftarrow v]}{t}$.
   \end{case}
@@
     $e = \textsf{if } v \: e_1 \: e_2$ and $e' = e_i$ where $i$ is determined by
-    the value of $v$.  \textsc{E-If}'s premises include $\typeRule{pc
-    \sqcup l}{e_i}{t}$ for $i \in \{1, 2\}$ and $l \guards t$, which follows
-    from Lemma~\ref{lem:weakening}.
+    the value of $v$.  \textsc{E-If}'s premises include $\typeRule{pc \sqcup
+    l}{e_i}{t}$ for $i \in \{1, 2\}$ and $l \guards t$, which follows from
+    Lemma~\ref{lem:weakening}.
   \end{case}
@@
     of \textsc{E-If} gives $l' \in H$ and $l' \guards t$.  Applying
-    Lemma~\ref{lem:projection} to the initial premises of \textsc{E-If} and using
-    the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup l}{e_i}$.
-    We can then apply \textsc{E-Bracket} to get the goal.
+    Lemma~\ref{lem:projection} to the initial premises of \textsc{E-If} and
+    using the result and applying \textsc{E-If} gives $\typeRule{pc \sqcup
+    l}{e_i}$.  We can then apply \textsc{E-Bracket} to get the goal.
   \end{case}
@@
   $v$ would be immediately uninteresting.  By \textsc{V-Bracket}, we know that
-  $\typeRule{ }{v}{t}$ holds given the assumptions.
-  By Lemma~\ref{lem:substitution}, we have $\typeRule{ }{e[x \Leftarrow
+  $\typeRule{ }{v}{t}$ holds given the assumptions.  By
+  Lemma~\ref{lem:substitution}, we have $\typeRule{ }{e[x \Leftarrow
   v]}{\textsf{int}^l}$.  By Lemma~\ref{lem:final_support}, we know that
@@
 As mentioned above, we unfortunately were unable to complete our noninterference
-proof in Coq by the time of this report.
-As of this writing, we have written definitions of a small-step operational semantics, a big-step operational semantics, typing judgments, and typing derivations for \langName.
-We have proven a few lemmas from section~\ref{sec:proof} relating to the operations semantics (including soundness), but remain mired in complexity on the remainder of the mechanization effort.
+proof in Coq by the time of this report.  As of this writing, we have written
+definitions of a small-step operational semantics, a big-step operational
+semantics, typing judgments, and typing derivations for \langName.  We have
+proven a few lemmas from section~\ref{sec:proof} relating to the operations
+semantics (including soundness), but remain mired in complexity on the remainder
+of the mechanization effort.
 
-In the process of mechanizing these lemmas, we empirically observed that the process of translating English into Coq can involve a nearly fifty-fold increase in the size of even the most basic proofs.
-While this was partially a natural effect of writing proofs in Coq, we found that much of the complexity stemmed from handling cases elided in the English-Language proofs; and in these cases we found hidden subtleties which appear to have escaped even Simonet and Pottier.
-We found ourselves
-often surprised by the number of items that were elided by the proofs in the
-paper.  For example, the proof of Lemma~\ref{lem:soundness} is one line of
-natural language but turned out to be 40 lines of Coq code, not including all of
-the supporting definitions and rules.
+In the process of mechanizing these lemmas, we empirically observed that the
+process of translating English into Coq can involve a nearly fifty-fold increase
+in the size of even the most basic proofs.  While this was partially a natural
+effect of writing proofs in Coq, we found that much of the complexity stemmed
+from handling cases elided in the English-Language proofs; and in these cases we
+found hidden subtleties which appear to have escaped even Simonet and Pottier.
+We found ourselves often surprised by the number of items that were elided by
+the proofs in the paper.  For example, the proof of Lemma~\ref{lem:soundness} is
+one line of natural language but turned out to be 40 lines of Coq code, not
+including all of the supporting definitions and rules.
 
@@
 
-We discovered a few small inaccuracies in \cite{InfoFlowML} during the 
-process of formalizing our proofs in Coq.  Listed below are the primary
-corrections we were able to make.
+We discovered a few small inaccuracies in \cite{InfoFlowML} during the process
+of formalizing our proofs in Coq.  Listed below are the primary corrections we
+were able to make.
 
-While some inaccuracies did invalidate Simonet and Pottier's original
-Lemmas, we universally found that a slightly different formulation was
-provable and remained usable in future proofs.
+While some inaccuracies did invalidate Simonet and Pottier's original Lemmas, we
+universally found that a slightly different formulation was provable and
+remained usable in future proofs.
 
@@
     As seen in Figure~\ref{fig:syntax}, we have a slightly different syntax
-    structure for \langName than was seen in~ \origLang.  Specifically, we constructed
-    values, answers, and expressions explicitly to not allow nesting of
-    bracketed values.  The syntax defined by Simonet and Pottier did allow
-    for this nesting, and authors dismissed is as a non-sensible
-    construction.  Unfortunately, allowing such syntactic nesting 
-    prevents the completeness Lemma from being proveable.
-    Realizing that our definitions did not disallow
+    structure for \langName than was seen in~ \origLang.  Specifically, we
+    constructed values, answers, and expressions explicitly to not allow nesting
+    of bracketed values.  The syntax defined by Simonet and Pottier did allow
+    for this nesting, and authors dismissed is as a non-sensible construction.
+    Unfortunately, allowing such syntactic nesting prevents the completeness
+    Lemma from being proveable.  Realizing that our definitions did not disallow
     this by construction after running into issues in the proof, we were able to
@@
     Proving our own version of Lemma~\ref{lem:soundness} was uniquely
-    instructive because it forced us to recognize that the original
-    formulation in Simonet and Pottier's paper on \origLang wasn't 
-    correct as stated.  Simonet and Pottier's Lemma asserts that 
-    if $e \rightarrow e'$, then $\lift{e}_i \rightarrow \lift{e'}_i$.
-    This does not hold in the case where $e$ is of the form $<v1 | v2> v$;
-    In this case, the single reduction rule that applies is \ref{eq:lift-app},
-    which does not step $\lift{e}_i \rightarrow \lift{e'}_i$ for any $i$ 
-    (as $\lift{e}_i$ and $\lift{e'}_i$ are identical).  Luckily, we are
-    fairly confident, given that we were still able to prove
-    Lemma~\ref{lem:stuck_bracket} despite this change, that this is not a
-    critical error in the original paper.
+    instructive because it forced us to recognize that the original formulation
+    in Simonet and Pottier's paper on \origLang wasn't correct as stated.
+    Simonet and Pottier's Lemma asserts that if $e \rightarrow e'$, then
+    $\lift{e}_i \rightarrow \lift{e'}_i$.  This does not hold in the case where
+    $e$ is of the form $<v1 | v2> v$; In this case, the single reduction rule
+    that applies is \ref{eq:lift-app}, which does not step $\lift{e}_i
+    \rightarrow \lift{e'}_i$ for any $i$ (as $\lift{e}_i$ and $\lift{e'}_i$ are
+    identical).  Luckily, we are fairly confident, given that we were still able
+    to prove Lemma~\ref{lem:stuck_bracket} despite this change, that this is not
+    a critical error in the original paper.
 \end{description}
@@
 
-Once our noninterference proof has been sucessfully mechanized for \langName, we will incrementally enhance the features of \langName, adding in absent constructs such as bind, references, and exceptions.
-We will then re-state and re-prove the noninterference result in a language very much like \origLang.
+Once our noninterference proof has been sucessfully mechanized for \langName, we
+will incrementally enhance the features of \langName, adding in absent
+constructs such as bind, references, and exceptions.  We will then re-state and
+re-prove the noninterference result in a language very much like \origLang.
 
@@
 us hope that it would be possible to implement something similar for a language
-like Jif.  We are aware that any such effort would require many person-years of effort to complete, but we remain convinced that such an effort would be a worthwhile initiative.
+like Jif.  We are aware that any such effort would require many person-years of
+effort to complete, but we remain convinced that such an effort would be a
+worthwhile initiative.
 
-With our verification effort complete, it would be possible to implement
-a compiler that produced proofs of noninterference to be carried around with the
+With our verification effort complete, it would be possible to implement a
+compiler that produced proofs of noninterference to be carried around with the
 compiled program.  This would be interesting in distributed settings where we

### 2013-12-19 paper/paper.tex
Typos and missing words.

@@
 checked and compiled by the system.  As we will discuss in
-Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to
+Section~\ref{sec:related}, these \emph{proof assistants} have been used to
 formalize, prove correct, and implement systems with verified properties.  This
@@
 
-Given the growing popularity of Proof Assistants information-flow languages, we
-were surprised to discover that there has not been work done to formalize ad
+Given the growing popularity of proof assistants and information-flow languages,
+we were surprised to discover that there has not been work done to formalize ad
 verify a language with a type system which enforces information flow security
@@
     Unfortunately, allowing such syntactic nesting prevents the completeness
-    Lemma from being proveable.  Realizing that our definitions did not disallow
+    Lemma from being provable.  Realizing that our definitions did not disallow
     this by construction after running into issues in the proof, we were able to
@@
 
-Once our noninterference proof has been sucessfully mechanized for \langName, we
-will incrementally enhance the features of \langName, adding in absent
+Once our noninterference proof has been successfully mechanized for \langName,
+we will incrementally enhance the features of \langName, adding in absent
 constructs such as bind, references, and exceptions.  We will then re-state and
